### 2022-09-19 docs/paper.tex
Add paper template

\documentclass[a4paper]{article}

\author{Jelmer Firet, Bram Pulles}
\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}

\begin{document}
\maketitle

\section{Pallets}

\section{Chip design}

\section{Dinner}

\section{Program safety}

\end{document}

### 2022-09-19 src/main.py

@@ -1,13 +1,0 @@
-from z3 import *
-
-x = Bool("x")
-y = Bool("y")
-
-solver = Solver()
-
-solver.add(And(x, Not(y)))
-solver.add(Not(y))
-
-print(solver.check())
-print("x =", solver.model()[x])
-print("y =", solver.model()[y])

### 2022-09-19 part1/chip.py
Chip readability

@@ -1,8 +1,10 @@
 from z3 import *
 import random
+import colorsys
 
 height = 30
 width = 30
+minDist = 17
 
 power = (3,4)
 numPower = 2
@@ -23,12 +25,8 @@ powerVars = []
 componentVars = []
 allVars = []
 
-solver = Optimize();
+solver = Solver();
 
-minDist = Int('minDist')
-solver.maximize(minDist)
-solver.add(16 <= minDist)
-solver.add(minDist <= 19)
 
 for i in range(numPower):
 	newVars = {};
@@ -69,25 +67,24 @@ for i in range(len(components)):
 	solver.add(0 <= newVars['top'])
 	solver.add(newVars['bottom'] <= height)
 	size = components[i]
-	solver.add(Or(And(newVars['right'] - newVars['left'] == size[0],
-	                  newVars['bottom'] - newVars['top'] == size[1]),
-	              And(newVars['right'] - newVars['left'] == size[1],
-	                 newVars['bottom'] - newVars['top'] == size[0])))
+	isVertical = And(newVars['right'] - newVars['left'] == size[0],
+	               newVars['bottom'] - newVars['top'] == size[1])
+	isHorizontal = And(newVars['right'] - newVars['left'] == size[1],
+	                 newVars['bottom'] - newVars['top'] == size[0])
+	solver.add(Or(isVertical, isHorizontal))
 
 	touches = []
 	for power in powerVars:
-		touches.append(And(power['left'] == newVars['right'],
-		                   power['top'] < newVars['bottom'],
-		                   power['bottom'] > newVars['top']))
-		touches.append(And(power['right'] == newVars['left'],
-		                   power['top'] < newVars['bottom'],
-		                   power['bottom'] > newVars['top']))
-		touches.append(And(power['top'] == newVars['bottom'],
-		                   power['left'] < newVars['right'],
-		                   power['right'] > newVars['left']))
-		touches.append(And(power['bottom'] == newVars['top'],
-		                   power['left'] < newVars['right'],
-		                   power['right'] > newVars['left']))
+		overlaps_y = And(power['top'] < newVars['bottom'],
+		                 power['bottom'] > newVars['top'])
+		touches.append(And(power['left'] == newVars['right'], overlaps_y))
+		touches.append(And(power['right'] == newVars['left'], overlaps_y))
+
[…]

### 2022-09-19 part1/pallets.py
Add intro

@@ -37,7 +37,7 @@ for truck in range(3,8):
 
 for truck in range(8):
 	solver.add(distribution['nuzzle'][truck] <= 1)
-	solver.add(Or(distribution['prittle'][truck] == 0, 
+	solver.add(Or(distribution['prittle'][truck] == 0,
 		          distribution['crottle'][truck] == 0))
 
 with open('pallets.smt', 'w') as file:
@@ -57,6 +57,3 @@ if (solver.check()):
 		print()
 else:
 	print('unsat')
-
-
-

### 2022-09-19 part1/pallets.py
Beautify pallets

@@ -1,59 +1,75 @@
 from z3 import *
 
-goods = {};
-goods['nuzzle'] = {'weight': 700, 'pallet': 4}
-goods['prittle'] = {'weight': 400}
-goods['skipple'] = {'weight': 1000, 'pallet': 8}
-goods['crottle'] = {'weight': 2500, 'pallet': 10}
-goods['dupple'] = {'weight': 200, 'pallet': 20}
+NUM_TRUCKS = 8
+TRUCK_CAPACITY = 8000
+TRUCK_MAX_PALLETS = 8
+
+# Create all the goods with their weight and number of pallets.
+goods = {
+    'nuzzle':  {'weight': 700,  'pallet': 4},
+    'skipple': {'weight': 1000, 'pallet': 8},
+    'crottle': {'weight': 2500, 'pallet': 10},
+    'dupple':  {'weight': 200,  'pallet': 20},
+    'prittle': {'weight': 400},
+};
+
+# Create all the z3 variables, an integer for every (good, truck) pair
+# representing the amount of that good in the truck.
 distribution = {}
 for good in goods.keys():
-	distribution[good] = [Int('{} {}'.format(good, truck)) for truck in range(8)]
-print(distribution)
+    distribution[good] = [Int('{} {}'.format(good, truck)) for truck in range(NUM_TRUCKS)]
 
+# Create the solver object, we optimize for the number of prittle pallets.
 solver = Optimize();
 
-for truck in range(8):
-	truckWeight = 0
-	numPallet = 0
-	for good in goods.keys():
-		truckWeight += distribution[good][truck] * goods[good]['weight']
-		numPallet += distribution[good][truck]
-		solver.add(0 <= distribution[good][truck])
-	solver.add(truckWeight <= 8000)
-	solver.add(numPallet <= 8)
+# Make sure that all the trucks do not have more than their max capacity of
+# pallets or more than their max capacity of weight.
+for truck in range(NUM_TRUCKS):
+    truckWeight = 0
+    numPallet = 0
 
+    for good in goods.keys():
+        truckWeight += distribution[good][truck] * goods[good]['weight']
+        numPallet += distribution[good][truck]
+        solver.add(0 <= distribution[good][truck])
+
+    solver.add(truckWeight <= TRUCK_CAPACITY)
+    solver.add(numPallet <= TRUCK_MAX_PALLETS)
+
+# Make sure that all the pallets for all the goods are distributed over the
+# trucks and optimise for the number of prittles.
 for good in goods.keys():
-	numPallet = 0
[…]

### 2022-09-19 docs/paper.tex
Add intro
Write pallets docs

@@
 \documentclass[a4paper]{article}
+
+\usepackage{amsfonts}
 
@@
 
+\section{Introduction}
+We have chosen to use Z3 for solving all of the assignments. In particular we
+use the pythonic Z3 library.
+
 \section{Pallets}
+In this task we need to distribute pallets of various goods over a limited
+number of trucks. To this end we have formulated the problem as a system of
+linear equations. We created an integer variable $v_{g,t}$ for every $good$ and
+$truck$ combination such that $v_{g,t} \in good \times truck$. This variable
+describes how many pallets of a certain good are in the specific truck. Using
+just these variables we can describe all of the constraints.
+
+\begin{itemize}
+	\item
+		We can only have a positive number of pallets in a truck.
+		\[ \forall_{v_{g,t} \in good \times truck}\ v_{g,t} \geq 0 \]
+	\item
+		Every truck has at most eight pallets.
+		\[ \forall_{t \in truck}\ \left( \sum_{g \in good} v_{g,t} \right) \leq 8 \]
+	\item
+		Every truck can carry at most 8000 kg. Let $w$ be a function giving the
+		weight of a given good $w: good \to \mathbb{N}$.
+		\[ \forall_{t \in truck}\ \left( \sum_{g \in good} w(g) \right) \leq 8000 \]
+	\item
+		For every good all the pallets are distributed. Except for prittles, as
+		no number of pallets is specified, we need to maximize this. Let $p$ be
+		a function giving the number of pallets of a given good $p: good \to
+		\mathbb{N}$.
+		\[
+		\forall_{g \in good \setminus prittles}\
+		\left( \sum_{t \in truck} v_{g,t} \right) = p(g)
+		\]
+	\item
+		Only three trucks can contain skipples. Let $t_i$ be the $i-$th truck,
+		starting at $i = 1$.
+		\[ \forall_{t_i \in truck, i > 3 }\ v_{skipples,t} = 0 \]
+	\item
+		No two pallets of nuzzles may be in the same truck.
+		\[ \forall_{t \in truck}\ v_{nuzzles,t} \leq 1 \]
+	\item
+		Prittles and crottles are not allowed to be put in the same truck.
+		\[ \forall_{t \in truck}\ v_{prittles,t} = 0 \lor v_{crottles,t} = 0 \]
+\end{itemize}
+
+All of the constraints described above can be easily converted to Z3. In order
+to let Z3 automatically maximize the number of pallets with prittles we set the
+maximisation function to maximize the formula shown below.
+
+\[ \sum_{t \in truck} v_{prittles,t} \]
+
+Running our program \textit{without} the last constraint (question 1) gives us
+the result shown below. Every row is a truck, every column is the good starting
+with that letter. As can be seen from the table, all the pallets can be
+distributed and there are a total of $8 + 5 + 5 + 4 = 22$ prittles pallets.
+\begin{verbatim}
+    n s c d p
+0 : 0 0 0 0 8
+1 : 0 8 0 0 0
+2 : 1 0 2 0 5
+3 : 1 0 2 0 5
+4 : 0 0 2 6 0
+5 : 0 0 0 4 4
+6 : 1 0 2 5 0
+7 : 1 0 2 5 0
+\end{verbatim}
+
+Running our program \textit{with} the last constraint (question 2) gives us the
+result shown below. Again, all the pallets can be distributed and there are a
+total of $4 + 8 + 8 = 20$ prittles pallets.
+
+\begin{verbatim}
+    n s c d p
+0 : 0 4 0 0 4
+1 : 1 2 2 1 0
+2 : 0 2 2 4 0
+3 : 0 0 0 0 8
+4 : 1 0 2 5 0
+5 : 1 0 2 5 0
+6 : 1 0 2 5 0
+7 : 0 0 0 0 8
+\end{verbatim}
+
+Our program and formalisation are generalised under the number of trucks, the
+truck maximum weight and maximum number of pallets. It is also very easy to add
+more goods, change their weight or change the number of pallets.
 

### 2022-09-19 docs/paper.tex
Add description of chips implementation

@@
-\documentclass[a4paper]{article}
+\documentclass{scrartcl}
+\usepackage{amsmath}
+\usepackage{fancyhdr}
+\usepackage{minted}
+\usepackage{enumitem}
+\usepackage{hyperref}
+\usepackage{svg}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{scalerel}
 
@@
 
-\author{Jelmer Firet, Bram Pulles}
+\author{Jelmer Firet (s1023433, \href{mailto:jelmer.firet@ru.nl}{jelmer.firet@ru.nl}) \and
+        Bram Pulles()}
 \title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}
+
+\pagestyle{fancy}
+\fancyhf{}
+\fancyhead[L]{Automated Reasoning PA 1|Jelmer Firet \& Bram Pulles}
+\fancyhead[R]{\thepage}
+
+\setminted{autogobble,tabsize=4,fontsize=\scriptsize}
+\setlength{\parindent}{0pt}
 
@@
 
+We consider the chip as a $30\times30$ grid.
+We number the rows from top to bottom, so the top row has index $0$ and the bottom row has index $29$.
+The columns are numbered the same way from left tot right.
+
+\subsection{variables} % (fold)
+\label{sub:chip_variables}
+	For each component $i$ we introduce $4$ integer variables: $\mathtt{left}_i$, $\mathtt{right}_i$, $\mathtt{top}_i$ and $\mathtt{bottom}_i$.\\
+	These variables indicate where the edges of component $i$ are.\\
+
+	\begin{tabular}{ll}
+		$\mathtt{left}_i$ & The left-most column that contains component $i$\\
+		$\mathtt{right}_i$ & The left-most column to the right of component $i$\\
+		$\mathtt{top}_i$ & The top-most column that contains component $i$\\
+		$\mathtt{bottom}_i$ & The top-most column below component $i$\\
+	\end{tabular}\\
+
+	Note that $\mathtt{left}_i$ is inclusive; the corresponding column contains component $i$.\\
+	However, $\mathtt{right}_i$ is exclusive; the corresponding column does not contain component $i$.\\
+	Inclusive-exclusive indexing ensures that $\mathtt{right}_i-\mathtt{left}_i$ is the width of component $i$.\\
+	We also use this indexing for the rows, so $\mathtt{bottom}_i-\mathtt{top}_i$ is the height of component $i$.\\
+	Component $s$ consists of the cells $(r,c)$ with $\mathtt{left}_i \leq c < \mathtt{right}_i$ and $\mathtt{top}_i \leq r < \mathtt{bottom}_i$
+
+	\begin{figure}[H]
+		\centering
+		\begin{subfigure}[b]{0.30\textwidth}
+			\centering
+			\includesvg[width=\textwidth]{chips-indexes}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.5\textwidth}
+			\centering
+			\begin{align*}
+				\mathtt{left}_i&=1 & \mathtt{right}_i &= 4\\
+				\mathtt{top}_i&=1 & \mathtt{bottom}_i &= 3
+			\end{align*}
+			\begin{align*}
+				\mathtt{width}_i=\mathtt{right}_i-\mathtt{left}_i&=4-1=3\\
+				\mathtt{height}_i=\mathtt{bottom}_i-\mathtt{top}_i&=3-1=2
+			\end{align*}
+		\end{subfigure}
+	\end{figure}
+% subsection chip_variables (end)
+
+\subsection{Constraints} % (fold)
+\label{sub:chip_constraints}
+\begin{enumerate}
+	\item 
+	We want every component to be placed inside the $30\times 30$ grid.\\
+	For this we add four constrains for each component:
+	\begin{align*}
+		0 &\leq \mathtt{left}_i & \mathtt{right}_i &\leq 30 & 0 &\leq \mathtt{top}_i & \mathtt{bottom}_i &\leq 30
+	\end{align*}
+	For every cell $(r,c)$ of component $i$ we then have:
+	\begin{align*}
+		0 \leq \mathtt{left}_i \leq c <  \mathtt{right}_i \leq 30 &&
+		0 \leq \mathtt{top}_i \leq r < \mathtt{bottom}_i \leq 30
+	\end{align*}
+	So each cell of the component is within the grid.
+
+	\item 
+	We want each component to have the correct size.\\
+	Suppose component $i$ should have size $(w,h)$ then we add the constraint:
+	\begin{align*}
+		(\mathtt{right}_i-\mathtt{left}_i = w\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = h)\;\lor\\
+		(\mathtt{right}_i-\mathtt{left}_i = h\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = w)
+	\end{align*}
+	The first line says that the component has the right size when it is not rotated.\\
+	The second line says that the component has the right size if it is rotated 90$^\circ$.\\
+	Because we want the component to have the right size in either orientation, we combine these clauses with an $\lor$.
+
+	\item 
+	We do not want any of the components to overlap.
+	For every pair of components $i<j$ we add the following constraint:
+	\begin{align*}
+		\mathtt{right}_i &\leq \mathtt{left}_j &\lor&&
+		\mathtt{left}_i &\geq \mathtt{right}_j &\lor&&
+		\mathtt{bottom}_i &\leq \mathtt{top}_j &\lor&&
+		\mathtt{top}_i &\geq \mathtt{bottom}_j
+	\end{align*}
+	This constrains say that component $j$ is entirely to the right, to the left, below or above component $i$. In each of these cases there is no overlap possible.
+
+	\item 
+	We want every regular component to share part of an edge with a power component.\\
+	For every regular component $i$ we add the following constraint:
+	\begin{align*}
+		\scalerel*{\lor}{\sum}_{p\in P}\left(
+		\begin{array}{l}
+			(\mathtt{right}_i = \mathtt{left}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
+			(\mathtt{left}_i = \mathtt{right}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
+			(\mathtt{bottom}_i = \mathtt{top}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)\;\lor\\
+			(\mathtt{top}_i = \mathtt{bottom}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)
+		\end{array}
+		\right)
+	\end{align*}
+
+	This statement starts with a big or over the power components. This ensures that there is a power component $p$ that touches component $i$.\\
+	Within the parenthesis we find four cases: component $i$ touches the left, the right, the top or the bottom of power component $p$.
+	To check if two components touch we need two parts.\\
+	First, both components need to touch a vertical or horizontal line from different sides. This is ensured by the first condition of every case.\\
+	Second, they need to share a part of that line, which is ensured by the other conditions.
+	We will only explain these other conditions for the first case.
+
+	In the first case the components touch a vertical line.
+	For the components to share a part of this line, there must be a row $r$ that contains both components.\\
+	This line must satisfy $\mathtt{top}_i \leq r < \mathtt{bottom}_i$ and $\mathtt{top}_p \leq r < \mathtt{bottom}_p$\\
+	We can combine these inequalities to form the following four conditions:
+	\begin{align*}
+		\mathtt{top}_i &< \mathtt{bottom}_i &\mathtt{top}_i &< \mathtt{bottom}_p & \mathtt{top}_p &< \mathtt{bottom}_i &\mathtt{top}_p &< \mathtt{bottom}_p
+	\end{align*}
+	The first and last condition are always satisfied (by part 2 and $0<w,h$).\\
+	The other conditions are added to the first case to guarantee overlap.
+
+	\item The centers of the power components are far enough apart
+	For every pair $p < q$ of power components we add the following constraint:
+	\begin{align*}
+		((\mathtt{top}_p + \mathtt{bottom}_p) - (\mathtt{top}_q + \mathtt{bottom}_q) &\geq 2\cdot D)\;\lor\\
+		((\mathtt{top}_q + \mathtt{bottom}_q) - (\mathtt{top}_p + \mathtt{bottom}_p) &\geq 2\cdot D)\;\lor\\
+		((\mathtt{left}_p + \mathtt{right}_p) - (\mathtt{left}_q + \mathtt{right}_q) &\geq 2\cdot D)\;\lor\\
+		((\mathtt{left}_q + \mathtt{right}_q) - (\mathtt{left}_p + \mathtt{right}_p) &\geq 2\cdot D)
+	\end{align*}
+
+	There are four cases: the center of component $p$ is $D$ units below, above, to the left or to the right of the center of component $q$.
+To ensure the vertical distance between the centers of two components we could use the following formula:
+	\begin{align*}
+		\lvert(\mathtt{top}_p + \mathtt{bottom}_p)/2 - (\mathtt{top}_q + \mathtt{bottom}_q)/2\rvert \geq D
+	\end{align*}
+	However we would like a formula without division or absolute values.
+	To get rid of the division we multiply everything by $2$.
+	We remove the absolute value by splitting the equation into two cases.
+	This equation then becomes the first two cases of our constraint.
+	The other two cases are found by considering the horizontal distance.
+\end{enumerate}
+% subsection chip_constraints (end)
+
+\subsection{Result} % (fold)
+\label{sub:result}
+The largest $D$ for which we can find a result is $17.5$.
+For $D=17.5$ it finds a result in $1.865$ seconds.
+For $D=18$ it proves that it is unsatisfiable in $67.782$ seconds.
+
+A solution for $D=17.5$ is illustrated below:
+\begin{figure}[H]
+	\centering
+	\includesvg[width=0.7\textwidth]{chips-solution}
+\end{figure}
+
+% subsection result (end)
+
 \section{Dinner}

### 2022-09-28 part1/chip.py
Solve dinner

@@ -4,7 +4,7 @@ import colorsys
 
 height = 30
 width = 30
-minDist = 17
+minDist = 17.5
 
 power = (3,4)
 numPower = 2

### 2022-09-28 part1/dinner.py

@@ -1,0 +1,109 @@
+from z3 import *
+
+numHouse = 5
+coupleSize = 2
+numPerson = coupleSize * numHouse
+numRound = 5
+groupPerRound = 2
+numGroup = groupPerRound * numRound
+# assume groupPerRound divides numPerson
+groupSize = numPerson // groupPerRound
+groupsPerHouse = numGroup // numHouse
+maxMeetup = 4
+
+
+inGroup = [[Bool('g{} p{}'.format(group, person))
+            for group in range(numGroup)]
+            for person in range(numPerson)]
+groupHost = [[Bool('g{} h{}'.format(group, house))
+              for house in range(numHouse)]
+              for group in range(numGroup)]
+
+solver = Solver()
+
+optional_a = []
+optional_b =[]
+optional_c = []
+optional_d = []
+optional_d2 = []
+
+for group in range(numGroup):
+	# At least 5 people in each group
+	visitors = [inGroup[person][group]
+	            for person in range(numPerson)]
+	solver.add(AtLeast(*visitors,groupSize))
+
+	for house in range(numHouse):
+		# The hosts are present in the group they host
+		hosts = [inGroup[person][group] for person in range(coupleSize*house, coupleSize*(house+1))]
+		solver.add(Implies(groupHost[group][house], And(*hosts)))
+
+		# If a couple does not host, they are not in the same group
+		optional_c.append(Implies(Not(groupHost[group][house]), AtMost(*hosts,1)))
+
+	# Each group is hosted by exactly one house
+	solver.add(AtLeast(*groupHost[group],1))
+	solver.add(AtMost(*groupHost[group],1))
+
+for person in range(numPerson):
+	# Each person visits at most one group each round
+	for round in range(numRound):
+		groups = inGroup[person][groupPerRound*round:groupPerRound*(round+1)]
+		solver.add(AtMost(*groups, 1))
+
+	for other in range(person+1, numPerson):
+		# Everyone meets each-other at most maxMeetup times
+		meetups = [And(inGroup[person][group], inGroup[other][group]) for group in range(numGroup)]
+		solver.add(AtMost(*meetups, maxMeetup))
+
+		# Every two people meet each-other at least once
+		optional_a.append(AtLeast(*meetups, 1))
+		# Every two people meet each-other at most three times
+		optional_b.append(AtMost(*meetups, 3))
+
+	for house in range(numHouse):
[…]

### 2022-09-28 part1/program.py
Add program safety

@@ -1,0 +1,58 @@
+from collections import namedtuple
+from z3 import *
+
+loops = 10
+
+Vars = namedtuple("Vars", list("abc"))
+v = Vars(
+        *[[Int("{}{}".format(v, i)) for i in range(loops + 1)] for v in "ab"],
+        [Bool("c{}".format(i)) for i in range(loops)]
+)
+
+solver = Solver()
+
+# initialise all the variables to 1
+solver.add(v.a[0] == 1)
+solver.add(v.b[0] == 1)
+
+# increase i with 1 every iteration
+for it in range(1, loops + 1):
+    solver.add(Implies(v.c[it - 1], And(
+        v.a[it] == v.a[it - 1] + 2 * v.b[it - 1],
+        v.b[it] == v.b[it - 1] + it
+    )))
+    solver.add(Implies(Not(v.c[it - 1]), And(
+        v.b[it] == v.a[it - 1] + v.b[it - 1],
+        v.a[it] == v.a[it - 1] + it
+    )))
+
+# simulate just to be sure of correctness
+def simulate(decisions, n):
+    a = 1
+    b = 1
+    for i in range(1, loops + 1):
+        if decisions[i - 1]:
+            a = a + 2 * b
+            b = b + i
+        else:
+            b = a + b
+            a = a + i
+    return b == 700 + n
+
+# check final condition
+for n in range(1, 11):
+    solver.push()
+
+    solver.add(v.b[loops] == 700 + n)
+    solvable = solver.check()
+
+    print("n =", n, "is", solvable, end="")
+    if solvable == sat:
+        print(" -> crash", end=" ")
+        decisions = [is_true(solver.model()[v.c[i]]) for i in range(loops)]
+        if simulate(decisions, n): print("verified")
+        else: print("WRONG")
+    else:
+        print()
+
+    solver.pop()

### 2022-10-10 part1/dinner.py
[Dinner] Better organisation of optional contraints

@@ -21,11 +21,7 @@ groupHost = [[Bool('g{} h{}'.format(group, house))
 
 solver = Solver()
 
-optional_a = []
-optional_b =[]
-optional_c = []
-optional_d = []
-optional_d2 = []
+optional = {'A': [], 'B': [], 'C': [], 'D': []}
 
 for group in range(numGroup):
 	# At least 5 people in each group
@@ -39,7 +35,7 @@ for group in range(numGroup):
 		solver.add(Implies(groupHost[group][house], And(*hosts)))
 
 		# If a couple does not host, they are not in the same group
-		optional_c.append(Implies(Not(groupHost[group][house]), AtMost(*hosts,1)))
+		optional['C'].append(Implies(Not(groupHost[group][house]), AtMost(*hosts,1)))
 
 	# Each group is hosted by exactly one house
 	solver.add(AtLeast(*groupHost[group],1))
@@ -57,16 +53,16 @@ for person in range(numPerson):
 		solver.add(AtMost(*meetups, maxMeetup))
 
 		# Every two people meet each-other at least once
-		optional_a.append(AtLeast(*meetups, 1))
+		optional['A'].append(AtLeast(*meetups, 1))
 		# Every two people meet each-other at most three times
-		optional_b.append(AtMost(*meetups, 3))
+		optional['B'].append(AtMost(*meetups, 3))
 
 	for house in range(numHouse):
 		if person // coupleSize == house:
 			continue
 		# Everyone is at most once a guest at another house
 		hosted = [And(inGroup[person][group], groupHost[group][house]) for group in range(numGroup)]
-		optional_d.append(AtMost(*hosted, 1))
+		optional['D'].append(AtMost(*hosted, 1))
 
 for house in range(numHouse):
 	# Everyone hosts at least groupsPerHouse times
@@ -91,16 +87,13 @@ subsets = [[]]
 for req in reqs:
 	subsets = [subset for subset in subsets] + [subset + [req] for subset in subsets]
 for subset in subsets:
-	solver.push()
 	print('with requirements {}'.format(', '.join(subset)))
-	if 'A' in subset:
-		solver.add(And(*optional_a))
-	if 'B' in subset:
-		solver.add(And(*optional_b))
-	if 'C' in subset:
-		solver.add(And(*optional_c))
-	if 'D' in subset:
-		solver.add(And(*optional_d))
+	if 'A' in subset and 'B' in subset:
+		print('unknown')
+		continue
[…]

### 2022-10-10 docs/paper.tex
Add mail
Add program verification explanation

@@
 \author{Jelmer Firet (s1023433, \href{mailto:jelmer.firet@ru.nl}{jelmer.firet@ru.nl}) \and
-        Bram Pulles()}
+Bram Pulles (s1015194, \href{mailto:bram.pulles@ru.nl}{bram.pulles@ru.nl})}
 \title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}
@@
 \section{Program safety}
+In order to test whether it is possible for the program to crash we use the SMT
+solver to check if the crashing statement can be reached. If the crash cannot
+be reached the formula will be unsatisfiable. If the crash can be reached the
+formula will be satisfiable and the variable assignment will provide us a value
+for the if-statement in every iteration.
+\\
+
+We start by creating a variable $l = 10$ for the number of iterations the loop
+will make. We also create $l + 1$ integer variables $a_i$ and $b_i$, where
+$a_0$ and $b_0$ are the initial values of $a$ and $b$, the others describe the
+values of $a$ and $b$ at every iteration of the loop. Lastly, we create $l$
+boolean $c_i$ variables to denote the value of the if-statement at iteration
+$i$.
+
+\begin{enumerate}
+	\item
+		We start with adding the constraints for the initial values of $a$ and $b$.
+		$$a_0 = 1 \land b_0 = 1$$
+
+	\item
+		If the if-statement is true we want to update the values of $a$ and $b$
+		accordingly. This is done using the following constraint.
+		\[
+			\forall_{1 \leq i \leq l + 1}
+			\left ( c_{i-1} \to
+				a_i = a_{i-1} + 2 \cdot b_{i-1} \land
+				b_i = b_{i-1} + i
+			\right )
+		\]
+	\item
+		If the if-statement is false we want to update the values of $a$ and $b$
+		accordingly. This is done using the following constraint.
+		\[
+			\forall_{1 \leq i \leq l + 1}
+			\left ( \lnot c_{i-1} \to
+				b_i = a_{i-1} + b_{i-1} \land
+				a_i = a_{i-1} + i
+			\right )
+		\]
+	\item
+		At last we want to check whether the program can crash, such that
+		$\text{SAT} \leftrightarrow \text{crash}$.
+		\[ b_l = 700 + n \]
+\end{enumerate}
+
+Running our program with all of these constraints yields the output shown
+below. The path shows the consecutive evaluations of the if-statement, with
+\texttt{f} as false and \texttt{t} as true, which leads to a crash.
+
+\begin{verbatim}
+n = 1 is sat -> crash
+path ffttttftff
+n = 2 is unsat
+n = 3 is unsat
+n = 4 is sat -> crash
+path ttfftftttf
+n = 5 is unsat
+n = 6 is sat -> crash
+path tftfffttff
+n = 7 is unsat
+n = 8 is sat -> crash
+path ttftftftft
+n = 9 is unsat
+n = 10 is sat -> crash
+path ttfttftttf
+\end{verbatim}
 

### 2022-10-10 part1/program.py

@@ -50,8 +50,11 @@ for n in range(1, 11):
     if solvable == sat:
         print(" -> crash", end=" ")
         decisions = [is_true(solver.model()[v.c[i]]) for i in range(loops)]
-        if simulate(decisions, n): print("verified")
-        else: print("WRONG")
+        if simulate(decisions, n):
+            print()
+            print("path", "".join(map(lambda b: "t" if b else "f", decisions)))
+        else:
+            print("WRONG")
     else:
         print()
 

### 2022-10-10 docs/paper.tex
[Dinner] write report

@@
 \usepackage{hyperref}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
 \usepackage{svg}
@@
 \usepackage{scalerel}
-
 \usepackage{amsfonts}
@@
 
+\subsection{Variables} % (fold)
+\label{sub:dinner_variables}
+We use the following terminology:
+\begin{itemize}
+	\item A \textbf{person} indicates a single person.
+	\item A \textbf{house} indicates a place where people dine together.
+	\item A \textbf{round} indicates a single day at which people dine together.
+	\item A \textbf{group} indicates a group that eats in the same house on the same day.
+\end{itemize}
+With $p\in h$ we will denote that person $p$ lives in house $h$.\\
+With $g\in r$ we will denote that group $g$ dines during round $r$.\\
+
+We use the following variables:\\
+\begin{tabular}{ll}
+	$\mathtt{inGroup}_{gp}$ & true when person $p$ dines with group $g$.\\
+	$\mathtt{groupHost}_{gh}$ & true when group $g$ dines at house $h$.\\
+\end{tabular}
+
+% subsection dinner_variables (end)
+
+\subsection{Constraints} % (fold)
+\label{sub:dinner_constraints}
+We have a lot of constraints that reference how many statements are true.\\
+We will use the following shorthand: $\sum_i S(i)$\\
+This denotes the number of true statements among the statements $S(i)$.
+\begin{itemize}
+	\item Due to the 1.5 meter restriction in every house at most 5 people may be present.\\
+	We add the following constraint for every group $g$:
+	$$\sum_{p}\mathtt{inGroup}_{gp}\leq 5$$
+	\item \ldots by which every round has to be prepared and served in two houses simultaneously, each with the corresponding couple and three guests.\\
+	This states, with a lot of words, that all $10$ people dine each round.\\
+	We add the following constraint for each round $r$ and person $p$:
+	$$\sum_{g\in r}\mathtt{inGroup}_{gp}\geq 1$$
+	\item \ldots each with the corresponding couple \ldots\\
+	This means that the owners of a house must be present at a dinner in their house.\\
+	We add the following constraint for each group $g$, house $h$ and person $p\in h$:
+	$$\mathtt{groupHost}_{gh}\to\mathtt{inGroup}_{gp}$$
+	Additionally it states that each group must be hosted by exactly one couple.
+	We add the following constraint for each group $g$:
+	$$\sum_h\mathtt{groupHost}_{gh} = 1$$
+	\item Every couple will serve two rounds in their house.\\
+	We add the following constraint for each house $h$:\\
+	$$\sum_g\mathtt{groupHost_{gh}}\geq 2$$
+	\item Every two people among the 10 participants meet each other at most 4 times during these 5 rounds.\\
+	We add the following constraint for every pair of persons $p_1<p_2$:\\
+	$$\sum_g(\mathtt{inGroup}_{gp_1}\land\mathtt{inGroup}_{gp_2})\leq 4$$
+\subsection{Optional constraints} % (fold)
+\label{sub:dinner_optional_constraints}
+\begin{enumerate}[label=\Alph*)]
+	\item Every two people among the 10 participants meet each other at least once.\\
+	We add the following constraint for every pair of persons $p_1<p_2$:
+	$$\sum_g(\mathtt{inGroup}_{gp_1}\land\mathtt{inGroup}_{gp_2})\geq 1$$
+	\item Every two people among the 10 participants meet each other at most $3$ times.\\
+	We add the following constraint for every pair of persons $p_1<p_2$:
+	$$\sum_g(\mathtt{inGroup}_{gp_1}\land\mathtt{inGroup}_{gp_2})\leq 3$$
+	\item Couples never meet outside their own houses\\
+	We add the following constraint for every group $g$ and house $h$:
+	$$\mathtt{groupHost}_{gh}\lor\sum_{p\in h}\mathtt{inGroup}_{pg}\leq 1$$
+	If a couple does not host a group there is at most one person of the couple in that group. Therefore they do not meet in that group.
+	\item For every house the six guests are distinct.\\
+	This is the case exactly when each house hosts each person at most once.
+	We add the following constraint for every house $h$ and person $p\notin h$:
+	$$\sum_{g}(\mathtt{groupHost}_{gh}\land \mathtt{inGroup}_{pg})\leq 1$$
+\end{enumerate}
+% subsection dinner_optional_constraints (end)
+
+\subsection{Results} % (fold)
+\label{sub:dinner_results}
+We have run Z3 for every subset of additional constraints with the standard constraints.
+We have noticed that Z3 takes too long to decide the cases where constraints A and B are both required.\\
+The only combination that was proven to be unsatisfiable was $A,C,D$. This implies that the combination of $A,B,C,D$ is also unsatisfiable.\\
+All remaining subsets are satisfiable.\\
+
+\begin{tabular}{ll|llll}
+\multicolumn{2}{l|}{} & & & A & A \\
+\multicolumn{2}{l|}{\multirow{-2}{*}{Constraints}} & & B & & B \\ \hline
+ & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{yellow!50}UNKNOWN \\
+ & D & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{yellow!50}UNKNOWN \\
+C & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{yellow!50}UNKNOWN \\
+C & D & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{red!50}UNSAT & \cellcolor{red!50}UNSAT 
+\end{tabular}
+
+Here are solutions to the bold subsets that imply satisfiability of all green subsets:\\
+
+\begin{tabular}{l|ll|lll|ll|lll}
+B,C,D & \multicolumn{5}{l|}{House 1} & \multicolumn{5}{l}{House 2} \\
+Round & \multicolumn{2}{l}{Host} & \multicolumn{3}{l|}{Guest} & \multicolumn{2}{l}{Host} & \multicolumn{3}{l}{Guest} \\ \hline
+1 & 0 & 1 & 3 & 5 & 9 & 6 & 7 & 2 & 4 & 8 \\
+2 & 0 & 1 & 2 & 4 & 6 & 8 & 9 & 3 & 5 & 7 \\
+3 & 2 & 3 & 0 & 6 & 8 & 4 & 5 & 1 & 7 & 9 \\
+4 & 2 & 3 & 1 & 7 & 9 & 4 & 5 & 0 & 6 & 8 \\
+5 & 6 & 7 & 1 & 3 & 5 & 8 & 9 & 0 & 2 & 4 \\
+\end{tabular}
+
+\begin{tabular}{l|ll|lll|ll|lll}
+A,C & \multicolumn{5}{l|}{House 1} & \multicolumn{5}{l}{House 2} \\
+Round & \multicolumn{2}{l}{Host} & \multicolumn{3}{l|}{Guest} & \multicolumn{2}{l}{Host} & \multicolumn{3}{l}{Guest} \\ \hline
+1 & 0 & 1 & 2 & 5 & 9 & 6 & 7 & 3 & 4 & 8 \\
+2 & 0 & 1 & 3 & 5 & 6 & 8 & 9 & 2 & 4 & 7 \\
+3 & 2 & 3 & 0 & 5 & 7 & 8 & 9 & 1 & 4 & 6 \\
+4 & 2 & 3 & 1 & 7 & 9 & 4 & 5 & 0 & 6 & 8 \\
+5 & 4 & 5 & 1 & 3 & 9 & 6 & 7 & 0 & 2 & 8 \\
+\end{tabular}
+
+\begin{tabular}{l|ll|lll|ll|lll}
+A,D & \multicolumn{5}{l|}{House 1} & \multicolumn{5}{l}{House 2} \\
+Round & \multicolumn{2}{l}{Host} & \multicolumn{3}{l|}{Guest} & \multicolumn{2}{l}{Host} & \multicolumn{3}{l}{Guest} \\ \hline
+1 & 0 & 1 & 6 & 7 & 9 & 2 & 3 & 4 & 5 & 8 \\
+2 & 0 & 1 & 2 & 4 & 8 & 6 & 7 & 3 & 5 & 9 \\
+3 & 2 & 3 & 0 & 1 & 6 & 4 & 5 & 7 & 8 & 9 \\
+4 & 4 & 5 & 0 & 1 & 3 & 8 & 9 & 2 & 6 & 7 \\
+5 & 6 & 7 & 0 & 2 & 4 & 8 & 9 & 1 & 3 & 5 \\
+\end{tabular}
+% subsection dinner_results (end)
+
+\end{itemize}
+% subsection dinner_constraints (end)
+
 \section{Program safety}

### 2022-10-10 part1/dinner.py

@@ -24,10 +24,10 @@ solver = Solver()
 optional = {'A': [], 'B': [], 'C': [], 'D': []}
 
 for group in range(numGroup):
-	# At least 5 people in each group
+	# At most 5 people in each group
 	visitors = [inGroup[person][group]
 	            for person in range(numPerson)]
-	solver.add(AtLeast(*visitors,groupSize))
+	solver.add(AtMost(*visitors,groupSize))
 
 	for house in range(numHouse):
 		# The hosts are present in the group they host
@@ -42,10 +42,10 @@ for group in range(numGroup):
 	solver.add(AtMost(*groupHost[group],1))
 
 for person in range(numPerson):
-	# Each person visits at most one group each round
+	# Each person visits at least one group each round
 	for round in range(numRound):
 		groups = inGroup[person][groupPerRound*round:groupPerRound*(round+1)]
-		solver.add(AtMost(*groups, 1))
+		solver.add(AtLeast(*groups, 1))
 
 	for other in range(person+1, numPerson):
 		# Everyone meets each-other at most maxMeetup times

### 2022-10-10 docs/paper.tex
[Dinner] Brute-force search

@@
 The only combination that was proven to be unsatisfiable was $A,C,D$. This implies that the combination of $A,B,C,D$ is also unsatisfiable.\\
+We have run a brute-force search using C++ on the rounds to show that A and B can not be both satisfied
 All remaining subsets are satisfiable.\\
@@
 \multicolumn{2}{l|}{\multirow{-2}{*}{Constraints}} & & B & & B \\ \hline
- & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{yellow!50}UNKNOWN \\
- & D & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{yellow!50}UNKNOWN \\
-C & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{yellow!50}UNKNOWN \\
+ & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{red!30}UNSAT \\
+ & D & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{red!30}UNSAT \\
+C & & \cellcolor{green!50}SAT & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{red!30}UNSAT \\
 C & D & \cellcolor{green!50}SAT & \cellcolor{green!50}\textbf{SAT} & \cellcolor{red!50}UNSAT & \cellcolor{red!50}UNSAT 

### 2022-10-10 part1/dinner.cpp

@@ -1,0 +1,82 @@
+#include <bits/stdc++.h>
+using namespace std;
+
+typedef vector<int> vi;
+typedef vector<vi> vvi;
+typedef vector<vvi> vvvi;
+typedef int64_t ll;
+typedef vector<ll> vl;
+typedef vector<vl> vvl;
+typedef pair<int,int> ii;
+typedef vector<ii> vii;
+typedef vector<vii> vvii;
+typedef queue<int> qi;
+
+int main() {
+	vi split = {0,0,0,0,0,1,1,1,1,1};
+	vvi splits;
+	vvvi contributes;
+	do {
+		if (split[0] == 1) break;
+		int numCouple = 0;
+		for (int i = 0; i < 5; i++) {
+			numCouple += split[2*i] == split[2*i+1];
+		}
+		if (numCouple < 2) continue;
+		vvi contributions(10, vi(10, 0));
+		for (int i = 0; i < 10; i++) {
+			for (int j = 0; j < 10; j++) {
+				contributions[i][j] += split[i] == split[j];
+			}
+		}
+		contributes.push_back(contributions);
+		splits.push_back(split);
+	} while (next_permutation(split.begin(), split.end()));
+	cout << contributes.size() << endl;
+	vi rounds(contributes.size(), 0);
+	for (int i = 0; i < 5; i++) {
+		rounds[i] = 1;
+	}
+
+	cout << "precomputation complete" << endl;
+
+	function<bool(int,int,vvi)> dfs = [&](int depth, int last, vvi meetup){
+		if (depth == 5) {
+			for (int i = 0; i < 10; i++) {
+				for (int j = i+1; j < 10; j++) {
+					if (meetup[i][j] < 1 or meetup[i][j] > 3) return false;
+				}
+			}
+			return true;
+		}
+		for (int round = last+1; round < (int) contributes.size(); round++) {
+			if (depth == 1) {
+				cout << round << endl;
+			}
+			bool isValid = true;
+			vvi meetup2 = meetup;
+			for (int i = 0; i < 10; i++) {
+				for (int j = i+1; j < 10; j++) {
+					meetup2[i][j] += contributes[round][i][j];
+					isValid &= meetup2[i][j] <= 3;
+				}
+			}
+			if (!isValid) continue;
+			if (dfs(depth+1, round, meetup2)) {
+				for (int i = 0; i < 10; i++) {
+					cout << splits[round][i] << " ";
[…]

### 2022-11-17 part2/report/paper.tex
Start part2 report

--- 2022-10-10 docs/paper.tex
@@
 \fancyhf{}
-\fancyhead[L]{Automated Reasoning PA 1|Jelmer Firet \& Bram Pulles}
+\fancyhead[L]{Automated Reasoning PA 2|Jelmer Firet \& Bram Pulles}
 \fancyhead[R]{\thepage}
@@
 
-\section{Introduction}
-We have chosen to use Z3 for solving all of the assignments. In particular we
-use the pythonic Z3 library.
-
-\section{Pallets}
-In this task we need to distribute pallets of various goods over a limited
-number of trucks. To this end we have formulated the problem as a system of
-linear equations. We created an integer variable $v_{g,t}$ for every $good$ and
-$truck$ combination such that $v_{g,t} \in good \times truck$. This variable
-describes how many pallets of a certain good are in the specific truck. Using
-just these variables we can describe all of the constraints.
-
-\begin{itemize}
-	\item
-		We can only have a positive number of pallets in a truck.
-		\[ \forall_{v_{g,t} \in good \times truck}\ v_{g,t} \geq 0 \]
-	\item
-		Every truck has at most eight pallets.
-		\[ \forall_{t \in truck}\ \left( \sum_{g \in good} v_{g,t} \right) \leq 8 \]
-	\item
-		Every truck can carry at most 8000 kg. Let $w$ be a function giving the
-		weight of a given good $w: good \to \mathbb{N}$.
-		\[ \forall_{t \in truck}\ \left( \sum_{g \in good} w(g) \right) \leq 8000 \]
-	\item
-		For every good all the pallets are distributed. Except for prittles, as
-		no number of pallets is specified, we need to maximize this. Let $p$ be
-		a function giving the number of pallets of a given good $p: good \to
-		\mathbb{N}$.
-		\[
-		\forall_{g \in good \setminus prittles}\
-		\left( \sum_{t \in truck} v_{g,t} \right) = p(g)
-		\]
-	\item
-		Only three trucks can contain skipples. Let $t_i$ be the $i-$th truck,
-		starting at $i = 1$.
-		\[ \forall_{t_i \in truck, i > 3 }\ v_{skipples,t} = 0 \]
-	\item
-		No two pallets of nuzzles may be in the same truck.
-		\[ \forall_{t \in truck}\ v_{nuzzles,t} \leq 1 \]
-	\item
-		Prittles and crottles are not allowed to be put in the same truck.
-		\[ \forall_{t \in truck}\ v_{prittles,t} = 0 \lor v_{crottles,t} = 0 \]
-\end{itemize}
-
-All of the constraints described above can be easily converted to Z3. In order
-to let Z3 automatically maximize the number of pallets with prittles we set the
-maximisation function to maximize the formula shown below.
-
-\[ \sum_{t \in truck} v_{prittles,t} \]
-
-Running our program \textit{without} the last constraint (question 1) gives us
-the result shown below. Every row is a truck, every column is the good starting
-with that letter. As can be seen from the table, all the pallets can be
-distributed and there are a total of $8 + 5 + 5 + 4 = 22$ prittles pallets.
-\begin{verbatim}
-    n s c d p
-0 : 0 0 0 0 8
-1 : 0 8 0 0 0
-2 : 1 0 2 0 5
-3 : 1 0 2 0 5
-4 : 0 0 2 6 0
-5 : 0 0 0 4 4
-6 : 1 0 2 5 0
-7 : 1 0 2 5 0
-\end{verbatim}
-
-Running our program \textit{with} the last constraint (question 2) gives us the
-result shown below. Again, all the pallets can be distributed and there are a
-total of $4 + 8 + 8 = 20$ prittles pallets.
-
-\begin{verbatim}
-    n s c d p
-0 : 0 4 0 0 4
-1 : 1 2 2 1 0
-2 : 0 2 2 4 0
-3 : 0 0 0 0 8
-4 : 1 0 2 5 0
-5 : 1 0 2 5 0
-6 : 1 0 2 5 0
-7 : 0 0 0 0 8
-\end{verbatim}
-
-Our program and formalisation are generalised under the number of trucks, the
-truck maximum weight and maximum number of pallets. It is also very easy to add
-more goods, change their weight or change the number of pallets.
-
-\section{Chip design}
-
-We consider the chip as a $30\times30$ grid.
-We number the rows from top to bottom, so the top row has index $0$ and the bottom row has index $29$.
-The columns are numbered the same way from left tot right.
-
-\subsection{variables} % (fold)
-\label{sub:chip_variables}
-	For each component $i$ we introduce $4$ integer variables: $\mathtt{left}_i$, $\mathtt{right}_i$, $\mathtt{top}_i$ and $\mathtt{bottom}_i$.\\
-	These variables indicate where the edges of component $i$ are.\\
-
-	\begin{tabular}{ll}
-		$\mathtt{left}_i$ & The left-most column that contains component $i$\\
-		$\mathtt{right}_i$ & The left-most column to the right of component $i$\\
-		$\mathtt{top}_i$ & The top-most column that contains component $i$\\
-		$\mathtt{bottom}_i$ & The top-most column below component $i$\\
-	\end{tabular}\\
-
-	Note that $\mathtt{left}_i$ is inclusive; the corresponding column contains component $i$.\\
-	However, $\mathtt{right}_i$ is exclusive; the corresponding column does not contain component $i$.\\
-	Inclusive-exclusive indexing ensures that $\mathtt{right}_i-\mathtt{left}_i$ is the width of component $i$.\\
-	We also use this indexing for the rows, so $\mathtt{bottom}_i-\mathtt{top}_i$ is the height of component $i$.\\
-	Component $s$ consists of the cells $(r,c)$ with $\mathtt{left}_i \leq c < \mathtt{right}_i$ and $\mathtt{top}_i \leq r < \mathtt{bottom}_i$
-
-	\begin{figure}[H]
-		\centering
-		\begin{subfigure}[b]{0.30\textwidth}
-			\centering
-			\includesvg[width=\textwidth]{chips-indexes}
-		\end{subfigure}
-		\begin{subfigure}[b]{0.5\textwidth}
-			\centering
-			\begin{align*}
-				\mathtt{left}_i&=1 & \mathtt{right}_i &= 4\\
-				\mathtt{top}_i&=1 & \mathtt{bottom}_i &= 3
-			\end{align*}
-			\begin{align*}
-				\mathtt{width}_i=\mathtt{right}_i-\mathtt{left}_i&=4-1=3\\
-				\mathtt{height}_i=\mathtt{bottom}_i-\mathtt{top}_i&=3-1=2
-			\end{align*}
-		\end{subfigure}
-	\end{figure}
-% subsection chip_variables (end)
-
-\subsection{Constraints} % (fold)
-\label{sub:chip_constraints}
-\begin{enumerate}
-	\item
-	We want every component to be placed inside the $30\times 30$ grid.\\
-	For this we add four constrains for each component:
-	\begin{align*}
-		0 &\leq \mathtt{left}_i & \mathtt{right}_i &\leq 30 & 0 &\leq \mathtt{top}_i & \mathtt{bottom}_i &\leq 30
-	\end{align*}
-	For every cell $(r,c)$ of component $i$ we then have:
-	\begin{align*}
-		0 \leq \mathtt{left}_i \leq c <  \mathtt{right}_i \leq 30 &&
-		0 \leq \mathtt{top}_i \leq r < \mathtt{bottom}_i \leq 30
-	\end{align*}
-	So each cell of the component is within the grid.
-
-	\item
-	We want each component to have the correct size.\\
-	Suppose component $i$ should have size $(w,h)$ then we add the constraint:
-	\begin{align*}
-		(\mathtt{right}_i-\mathtt{left}_i = w\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = h)\;\lor\\
-		(\mathtt{right}_i-\mathtt{left}_i = h\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = w)
-	\end{align*}
-	The first line says that the component has the right size when it is not rotated.\\
-	The second line says that the component has the right size if it is rotated 90$^\circ$.\\
-	Because we want the component to have the right size in either orientation, we combine these clauses with an $\lor$.
-
-	\item
-	We do not want any of the components to overlap.
-	For every pair of components $i<j$ we add the following constraint:
-	\begin{align*}
-		\mathtt{right}_i &\leq \mathtt{left}_j &\lor&&
-		\mathtt{left}_i &\geq \mathtt{right}_j &\lor&&
-		\mathtt{bottom}_i &\leq \mathtt{top}_j &\lor&&
-		\mathtt{top}_i &\geq \mathtt{bottom}_j
-	\end{align*}
-	This constrains say that component $j$ is entirely to the right, to the left, below or above component $i$. In each of these cases there is no overlap possible.
-
-	\item
-	We want every regular component to share part of an edge with a power component.\\
-	For every regular component $i$ we add the following constraint:
-	\begin{align*}
-		\scalerel*{\lor}{\sum}_{p\in P}\left(
-		\begin{array}{l}
-			(\mathtt{right}_i = \mathtt{left}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
-			(\mathtt{left}_i = \mathtt{right}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
-			(\mathtt{bottom}_i = \mathtt{top}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)\;\lor\\
-			(\mathtt{top}_i = \mathtt{bottom}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)
-		\end{array}
-		\right)
-	\end{align*}
-
-	This statement starts with a big or over the power components. This ensures that there is a power component $p$ that touches component $i$.\\
-	Within the parenthesis we find four cases: component $i$ touches the left, the right, the top or the bottom of power component $p$.
-	To check if two components touch we need two parts.\\
-	First, both components need to touch a vertical or horizontal line from different sides. This is ensured by the first condition of every case.\\
-	Second, they need to share a part of that line, which is ensured by the other conditions.
-	We will only explain these other conditions for the first case.
-
-	In the first case the components touch a vertical line.
-	For the components to share a part of this line, there must be a row $r$ that contains both components.\\
-	This line must satisfy $\mathtt{top}_i \leq r < \mathtt{bottom}_i$ and $\mathtt{top}_p \leq r < \mathtt{bottom}_p$\\
-	We can combine these inequalities to form the following four conditions:
-	\begin{align*}
-		\mathtt{top}_i &< \mathtt{bottom}_i &\mathtt{top}_i &< \mathtt{bottom}_p & \mathtt{top}_p &< \mathtt{bottom}_i &\mathtt{top}_p &< \mathtt{bottom}_p
-	\end{align*}
-	The first and last condition are always satisfied (by part 2 and $0<w,h$).\\
-	The other conditions are added to the first case to guarantee overlap.
-
-	\item The centers of the power components are far enough apart
-	For every pair $p < q$ of power components we add the following constraint:
-	\begin{align*}
-		((\mathtt{top}_p + \mathtt{bottom}_p) - (\mathtt{top}_q + \mathtt{bottom}_q) &\geq 2\cdot D)\;\lor\\
-		((\mathtt{top}_q + \mathtt{bottom}_q) - (\mathtt{top}_p + \mathtt{bottom}_p) &\geq 2\cdot D)\;\lor\\
-		((\mathtt{left}_p + \mathtt{right}_p) - (\mathtt{left}_q + \mathtt{right}_q) &\geq 2\cdot D)\;\lor\\
-		((\mathtt{left}_q + \mathtt{right}_q) - (\mathtt{left}_p + \mathtt{right}_p) &\geq 2\cdot D)
-	\end{align*}
-
-	There are four cases: the center of component $p$ is $D$ units below, above, to the left or to the right of the center of component $q$.
-To ensure the vertical distance between the centers of two components we could use the following formula:
-	\begin{align*}
-		\lvert(\mathtt{top}_p + \mathtt{bottom}_p)/2 - (\mathtt{top}_q + \mathtt{bottom}_q)/2\rvert \geq D
-	\end{align*}
-	However we would like a formula without division or absolute values.
-	To get rid of the division we multiply everything by $2$.
-	We remove the absolute value by splitting the equation into two cases.
-	This equation then becomes the first two cases of our constraint.
-	The other two cases are found by considering the horizontal distance.
-\end{enumerate}
-% subsection chip_constraints (end)
-
-\subsection{Result} % (fold)
-\label{sub:result}
-The largest $D$ for which we can find a result is $17.5$.
-For $D=17.5$ it finds a result in $1.865$ seconds.
-For $D=18$ it proves that it is unsatisfiable in $67.782$ seconds.
-
-A solution for $D=17.5$ is illustrated below:
-\begin{figure}[H]
-	\centering
-	\includesvg[width=0.7\textwidth]{chips-solution}
-\end{figure}
-
-% subsection result (end)
-
-\section{Dinner}
-
-\section{Program safety}
-In order to test whether it is possible for the program to crash we use the SMT
-solver to check if the crashing statement can be reached. If the crash cannot
-be reached the formula will be unsatisfiable. If the crash can be reached the
-formula will be satisfiable and the variable assignment will provide us a value
-for the if-statement in every iteration.
-\\
-
-We start by creating a variable $l = 10$ for the number of iterations the loop
-will make. We also create $l + 1$ integer variables $a_i$ and $b_i$, where
-$a_0$ and $b_0$ are the initial values of $a$ and $b$, the others describe the
-values of $a$ and $b$ at every iteration of the loop. Lastly, we create $l$
-boolean $c_i$ variables to denote the value of the if-statement at iteration
-$i$.
-
-\begin{enumerate}
-	\item
-		We start with adding the constraints for the initial values of $a$ and $b$.
-		$$a_0 = 1 \land b_0 = 1$$
-
-	\item
-		If the if-statement is true we want to update the values of $a$ and $b$
-		accordingly. This is done using the following constraint.
-		\[
-			\forall_{1 \leq i \leq l + 1}
-			\left ( c_{i-1} \to
-				a_i = a_{i-1} + 2 \cdot b_{i-1} \land
-				b_i = b_{i-1} + i
-			\right )
-		\]
-	\item
-		If the if-statement is false we want to update the values of $a$ and $b$
-		accordingly. This is done using the following constraint.
-		\[
-			\forall_{1 \leq i \leq l + 1}
-			\left ( \lnot c_{i-1} \to
-				b_i = a_{i-1} + b_{i-1} \land
-				a_i = a_{i-1} + i
-			\right )
-		\]
-	\item
-		At last we want to check whether the program can crash, such that
-		$\text{SAT} \leftrightarrow \text{crash}$.
-		\[ b_l = 700 + n \]
-\end{enumerate}
-
-Running our program with all of these constraints yields the output shown
-below. The path shows the consecutive evaluations of the if-statement, with
-\texttt{f} as false and \texttt{t} as true, which leads to a crash.
-
-\begin{verbatim}
-n = 1 is sat -> crash
-path ffttttftff
-n = 2 is unsat
-n = 3 is unsat
-n = 4 is sat -> crash
-path ttfftftttf
-n = 5 is unsat
-n = 6 is sat -> crash
-path tftfffttff
-n = 7 is unsat
-n = 8 is sat -> crash
-path ttftftftft
-n = 9 is unsat
-n = 10 is sat -> crash
-path ttfttftttf
-\end{verbatim}
-
 \end{document}

### 2022-11-17 part2/report/paper.tex
Report and robot

@@
 
+\section{Groups} % (fold)
+\label{sec:groups}
+\begin{enumerate}[label=\alph*)]
+\item 
+	Using prover9 we can prove that $I*x=x$, $inv(inv(x))=x$ and $inv(x)*x=I$.\\
+	However, it is not true in general that $inv(x*y)=inv(x)*inv(y)$.\\
+	The smallest group for which this does not hold has size $6$.
+\item 
+	The smallest non-Abelian group has size $6$.
+\item 
+	For $n=2$ all such groups are Abelian.\\
+	For $n=3$ the group is not Abelian in general and the smallest counterexample has size $27$.\\
+	For $n=4$ the group is not Abelian and the smallest counterexample has size $8$.
+\end{enumerate}
+% section groups (end)
+
 \end{document}

### 2022-11-17 part2/robot.py

@@ -1,0 +1,118 @@
+from z3 import *
+import csv
+
+START = 0
+TARGET = 1
+LAVA = 2
+ICE = 3
+
+NORTH = 0
+EAST = 1
+SOUTH = 2
+WEST = 3
+
+optimize = Optimize()
+goal = Int('goal')
+optimize.minimize(goal)
+
+with open('grid3.csv', mode='r') as file:
+	csvFile = csv.reader(file)
+	grid = [[int(x) for x in line] for line in csvFile]
+
+height = len(grid)
+width = len(grid[0])
+numColor = 0
+for r in range(height):
+	for c in range(width):
+		numColor = max(grid[r][c]+1, numColor)
+
+direction = [Int("dir {}".format(i)) for i in range(numColor)]
+for i in range(numColor):
+	optimize.add(0 <= direction[i])
+	optimize.add(direction[i] < 4)
+
+def next(r,c,direction, dist):
+	if direction == NORTH:
+		return max(r-dist, 0), c
+	if direction == EAST:
+		return r, min(c+dist, width-1)
+	if direction == SOUTH:
+		return min(r+dist, height-1), c
+	if direction == WEST:
+		return r, max(c-dist, 0)
+
+
+dist = [[Int("dist {} {}".format(r,c)) for c in range(width)] for r in range(height)]
+for r in range(height):
+	for c in range(width):
+		color = grid[r][c]
+		if color == TARGET:
+			optimize.add(dist[r][c] == 0)
+		elif color == LAVA:
+			optimize.add(dist[r][c] == goal+1) # practically infinite
+		elif color == ICE:
+			for dir in range(4):
+				next_r, next_c = next(r,c,dir,1)
+				optimize.add(Implies(direction[color] == dir,
+					                 Or(dist[r][c] == goal+1, # practically infinite
+					                 	dist[r][c] >= dist[next_r][next_c]+1
+				)))
+
+				next_r, next_c = next(r,c,dir,2)
+				optimize.add(Implies(direction[color] == dir,
+					                 Or(dist[r][c] == goal+1, # practically infinite
+					                 	dist[r][c] >= dist[next_r][next_c]+1
+				)))
+
+		else:
+			if color == START:
+				optimize.add(dist[r][c] <= goal)
+			for dir in range(4):
[…]

### 2022-11-21 part2/report/paper.tex
Add groups docs

--- 2022-11-17 part2/report/paper.tex
@@
 \documentclass{scrartcl}
+
 \usepackage{amsmath}
@@
 Bram Pulles (s1015194, \href{mailto:bram.pulles@ru.nl}{bram.pulles@ru.nl})}
-\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}
+\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 2}}
 
@@
 
+\section{Groups} % (fold)
+\label{sec:groups}
+\begin{enumerate}[label=\alph*)]
+\item
+	We proved that $I*x=x$, $inv(inv(x))=x$ and $inv(x)*x=I$ by putting all of
+	the given formulas \textit{literally} in prover9, see
+	\texttt{groups\_a1.in}.
+
+	Using mace4, we show that, in general, $inv(x*y)=inv(x)*inv(y)$ does not
+	hold. The smallest group for which this does not hold has size $6$, see
+	\texttt{groups\_a2.in} and appendix \ref{apx: groups}.
+\item
+	Using mace4, we show that the smallest non-Abelian group has size $6$,
+	again by \textit{literally} putting in the formulas provided, see
+	\texttt{groups\_b.in} and appendix \ref{apx: groups}.
+\item
+	We use the fact that $x^2 = x * x$, $x^3 = x * (x * x)$ and $x^4 = x * (x *
+	(x * x))$ this makes encoding $x^n = I$ straightforward for $n = 2, 3, 4$.
+
+	Using prover9, we show that for $n=2$ all such groups are Abelian, see \texttt{groups\_c1.in}.
+
+	Using mace4, we show that for $n=3$ the group is not Abelian in general and
+	the smallest counterexample has size $27$, see \texttt{groups\_c2.in}.
+
+	Using mace4, we show that for $n=4$ the group is not Abelian and the
+	smallest counterexample has size $8$, see \texttt{groups\_c3.in} and
+	appendix \ref{apx: groups}.
+\end{enumerate}
+% section groups (end)
+
+\appendix
+\pagebreak
+\section{Groups}
+\label{apx: groups}
+\begin{enumerate}[label=\alph*)]
+	\item
+		Smallest finite group for which $inv(x*y)=inv(x)*inv(y)$ does not hold.
+		\begin{verbatim}
+		interpretation( 6, [number=1, seconds=0], [
+
+		    function(I, [ 0 ]),
+
+		    function(c1, [ 1 ]),
+
+		    function(c2, [ 2 ]),
+
+		    function(inv(_), [ 0, 1, 2, 4, 3, 5 ]),
+
+		    function(*(_,_), [
+		        0, 1, 2, 3, 4, 5,
+		        1, 0, 3, 2, 5, 4,
+		        2, 4, 0, 5, 1, 3,
+		        3, 5, 1, 4, 0, 2,
+		        4, 2, 5, 0, 3, 1,
+		        5, 3, 4, 1, 2, 0 ])
+		]).
+		\end{verbatim}
+
+	\item
+		Smallest non-abelian group.
+		\begin{verbatim}
+		interpretation( 6, [number=1, seconds=0], [
+
+			    function(I, [ 0 ]),
+
+			    function(c1, [ 1 ]),
+
+			    function(c2, [ 2 ]),
+
+			    function(inv(_), [ 0, 1, 2, 4, 3, 5 ]),
+
+			    function(*(_,_), [
+			        0, 1, 2, 3, 4, 5,
+			        1, 0, 3, 2, 5, 4,
+			        2, 4, 0, 5, 1, 3,
+			        3, 5, 1, 4, 0, 2,
+			        4, 2, 5, 0, 3, 1,
+			        5, 3, 4, 1, 2, 0 ])
+		]).
+		\end{verbatim}
+
+	\item
+		\pagebreak
+		Smallest non-abelian group with $x^4 = I$.
+		\begin{verbatim}
+		interpretation( 6, [number=1, seconds=0], [
+
+			    function(I, [ 0 ]),
+
+			    function(c1, [ 1 ]),
+
+			    function(c2, [ 2 ]),
+
+			    function(inv(_), [ 0, 1, 2, 4, 3, 5, 6, 7 ]),
+
+			    function(*(_,_), [
+			        0, 1, 2, 3, 4, 5, 6, 7,
+			        1, 0, 3, 2, 5, 4, 7, 6,
+			        2, 4, 0, 6, 1, 7, 3, 5,
+			        3, 5, 1, 7, 0, 6, 2, 4,
+			        4, 2, 6, 0, 7, 1, 5, 3,
+			        5, 3, 7, 1, 6, 0, 4, 2,
+			        6, 7, 4, 5, 2, 3, 0, 1,
+			        7, 6, 5, 4, 3, 2, 1, 0 ])
+		]).
+		\end{verbatim}
+\end{enumerate}
+
 \end{document}

### 2022-11-21 part2/report/paper.tex
Write report for robot

@@
 
+\section{Robot} % (fold)
+\label{sec:robot}
+
+	\subsection{Explanation} % (fold)
+	\label{sub:explanation}
+	In this problem we are given a grid with colored cells. We should program a robot for this grid so that the robot will always find it's way to one of the goal cells if it starts on any start cell.
+	The robot can detect the color of the cell it is on and decide based on that color which direction to move.
+Our program should find the best assignment of directions to colors.
+We will Z3 to find this assignment.
+
+	In our program we read the grid from a csv file.
+	We then determine the number of colors in this grid.
+	For each color we make a Z3 integer that will store the direction to move for that color. The values of this integer correspond to the following directions:\\
+
+	\begin{tabular}{l|ll}
+		value & direction & effect on coordinates\\
+		\hline
+		0 & north & decreasing row\\
+		1 & east & increasing column\\
+		2 & south & increasing row\\
+		3 & west & decreasing column\\
+	\end{tabular}\\
+
+	To ensure that Z3 only assigns one of these, we add the following constraint:
+	$$\mathtt{\forall_{color}: 0 \le direction[color] < 4}$$
+
+	Now we want to evaluate the performance of an assignment.
+For this we want to know for every cell if we reach a destination, and if so in how many steps.
+We will add a Z3 integer storing this distance for every cell.
+For now we will say that this integer has the value $\infty$ if it can not reach a destination cell.\\
+
+	We then specify how the distances interact with the commands.
+We will start with the simple cases; the destination and the lava.
+For a destination cell we do not need to move to reach a destination cell, therefore we add the following constraint:
+	$$\mathtt{\forall_{0\le r<R}\forall_{0\le c<C}\text{ if }grid[r][c]=DESTINATION: dist[r][c]=0}$$
+	We do something similar for lava. From a lava cell we cannot reach a destination cell, so we add the following constraint:
+	$$\mathtt{\forall_{0\le r<R}\forall_{0\le c<C}\text{ if }grid[r][c]=LAVA: dist[r][c]=\infty}$$
+
+	Now we will specify what the robot does for regular cells.
+For convenience we have a function $\mathtt{next(r,c,direction,steps)}$ that give the position we go to when we take $\mathtt{steps}$ steps to $\mathtt{direction}$ from cell $(\mathtt{r},\mathtt{c})$.
+This function takes into account that the robot stops at the edges of the grid.
+	\begin{align*}
+		&\mathtt{\forall_{0\le r<R}\forall_{0\le c<C}\forall_{0\le dir<4}\text{ if }\neg special(grid[r][c]):}\\
+		&\mathtt{direction[grid[r][c]]=dir \implies dist[r][c]=1+dist[nr][nc]}\\
+		&\text{ where }
+		\mathtt{(nr,nc) = next(r,c,dir,1)}
+	\end{align*}
+	Note that we consider $1+\infty=\infty$. When we give $\infty$ an actual value, we will have to specify this special case explicitly.
+
+	We can now specify how the ice cells work.
+We want $\mathtt{dist[r][c]}$ to store the worst-case distance to a destination cell.
+We will have a similar condition to the one for the regular cells, but take the maximum of the cells where we can end up.
+This translates into the following constraint:
+	\begin{align*}
+		&\mathtt{\forall_{0\le r<R}\forall_{0\le c<C}\forall_{0\le dir<4}\text{ if } grid[r][c] = ICE:}\\
+		&\mathtt{direction[grid[r][c]]=dir \implies dist[r][c]=1+\max(dist[nr][nc],dist[nr',nc'])}\\
+		&\text{ where }
+		\mathtt{(nr,nc) = next(r,c,dir,1) \text{ and } (nr',nc') = next(r,c,dir,2)}
+	\end{align*}
+
+	Now we are ready to specify what we want to achieve, we want an assignment of directions to colors that minimizes
+	$$\mathtt{\max_{0\le r < R, 0\le c < C, grid[r][c]=START} dist[r][c]}$$
+	We do this by adding a Z3 integer for this goal and ask the Z3 optimizer to minimize this variable. Then we add the following constraint:
+	$$\mathtt{\forall_{0\le r<R}\forall_{0\le c<C}\text{ if }grid[r][c]=START: dist[r][c]\le goal}$$
+
+	Now we can choose a value for $\infty$.
+We do not want the $\mathtt{dist[r][c]}$ of start cells to be infinite, but otherwise the value does not matter.
+Therefore we choose the value of $\infty$ to be $\mathtt{goal+1}$.
+Then in all of the previous constraints of the form $\mathtt{dist[r][c]=\ldots}$ we add the extra option that $\mathtt{dist[r][c]=goal+1}$.
+This allows Z3 to assign values to all distances even in the case of an infinite loop.
+	% subsection explanation (end)
+
+	\subsection{Results} % (fold)
+	\label{sub:results}
+	When we run our program on the demogrid, we find that the minimal solution uses 11 steps\footnote{The problem statement gives a "solution for 10 steps", but the image shows a solution with 11 steps}.
+The assignment of directions is visualized in figure \ref{robot-demogrid-solution}.
+
+	On grid (a) our program finds that the minimal solution uses 14 steps. The assignment of directions is visualized in figure \ref{robot-grid-a-solution}
+
+	On grid (b) our program also finds a minimal solution of 14 steps, which is visualised in figure \ref{robot-grid-b-solution}
+
+	On grid (c) our program also finds a minimal solution of 18 steps, which is visualised in figure \ref{robot-grid-c-solution}
+
+	\begin{figure}
+		\begin{subfigure}{0.5\textwidth}
+			\includegraphics[width=0.9\textwidth]{resources/Screenshot from 2022-11-21 12-14-49.png}
+			\caption{Solution for demo grid}
+			\label{robot-demogrid-solution}
+		\end{subfigure}
+		\begin{subfigure}{0.5\textwidth}
+			\includegraphics[width=0.9\textwidth]{resources/Screenshot from 2022-11-21 12-20-11.png}
+			\caption{Solution for grid (a)}
+			\label{robot-grid-a-solution}
+		\end{subfigure}
+		\begin{subfigure}{0.5\textwidth}
+			\includegraphics[width=0.9\textwidth]{resources/Screenshot from 2022-11-21 12-21-43.png}
+			\caption{Solution for grid (b)}
+			\label{robot-grid-b-solution}
+		\end{subfigure}
+		\begin{subfigure}{0.5\textwidth}
+			\includegraphics[width=0.9\textwidth]{resources/Screenshot from 2022-11-21 12-23-17.png}
+			\caption{Solution for grid (c)}
+			\label{robot-grid-c-solution}
+		\end{subfigure}
+
+	\end{figure}
+	% subsection results (end)
+
+% section robot (end)
 \appendix

### 2022-11-21 part2/robot.py

@@ -31,15 +31,15 @@ for i in range(numColor):
 	optimize.add(0 <= direction[i])
 	optimize.add(direction[i] < 4)
 
-def next(r,c,direction, dist):
+def next(r,c,direction, steps):
 	if direction == NORTH:
-		return max(r-dist, 0), c
+		return max(r-steps, 0), c
 	if direction == EAST:
-		return r, min(c+dist, width-1)
+		return r, min(c+steps, width-1)
 	if direction == SOUTH:
-		return min(r+dist, height-1), c
+		return min(r+steps, height-1), c
 	if direction == WEST:
-		return r, max(c-dist, 0)
+		return r, max(c-steps, 0)
 
 
 dist = [[Int("dist {} {}".format(r,c)) for c in range(width)] for r in range(height)]

### 2022-12-15 part2/multiset-path-order.py
Update mpo

@@ -5,7 +5,7 @@ import sys
 
 class Term:
 	""" The base class for terms """
-	def toString(self) -> str:
+	def toString(self, print_index=True) -> str:
 		pass
 
 	def print(self) -> None:
@@ -42,8 +42,8 @@ class Function(Term):
 		self.name = name
 		self.args = args
 
-	def toString(self) -> str:
-		argStrings = [arg.toString() for arg in self.args]
+	def toString(self, print_index=True) -> str:
+		argStrings = [arg.toString(print_index) for arg in self.args]
 		return self.name + '(' + ', '.join(argStrings) + ')'
 	
 	def subterms(self):
@@ -63,8 +63,11 @@ class Variable(Term):
 		self.name = name
 		self.index = index
 
-	def toString(self) -> str:
-		return self.name + '{' + str(self.index) + '}'
+	def toString(self, print_index=True) -> str:
+		if print_index:
+			return self.name + '{' + str(self.index) + '}'
+		else:
+			return self.name
 
 	def subterms(self):
 		return [self]
@@ -109,13 +112,13 @@ def parse_inequality(string : str, index : int) -> Comparison:
 	return Comparison(lhs, rhs)
 
 inequalities = []
-# inequalities.append(parse_inequality('c(x,y,u,v) > b(f(x,y),b(u,u,u),g(v,b(x,y,u)))',1))
-# inequalities.append(parse_inequality('b(f(x,y),g(x,y),f(x,g(z,u))) > b(f(x,z),y,g(g(g(y,x),x),x))',2))
-# inequalities.append(parse_inequality('h(g(x,g(u,z)),c(x,y,x,z)) > f(d(x,z),u)',3))
-# inequalities.append(parse_inequality('h(d(f(x,y),g(u,v)),f(x,y)) > f(c(u,x,v,y),g(y,x))',4))
-# inequalities.append(parse_inequality('f(b(x,y,z),u) > h(u,f(x,h(y,x)))',5))
-# inequalities.append(parse_inequality('b(a(x,y,z),y,x) > c(x,x,y,x)',6))
-inequalities.append(parse_inequality('f(x) > x',7))
+inequalities.append(parse_inequality('c(x,y,u,v) > b(f(x,y),b(u,u,u),g(v,b(x,y,u)))',1))
+inequalities.append(parse_inequality('b(f(x,y),g(x,y),f(x,g(z,u))) > b(f(x,z),y,g(g(g(y,x),x),x))',2))
[…]

### 2022-12-15 part2/report/paper.tex
Report on mosaic

@@
 \setlength{\parindent}{0pt}
+
+\newcommand{\black}{\cellcolor{black}\color{white}}
 
@@
 % section robot (end)
+
+\pagebreak
+\section{Mosaic} % (fold)
+\label{sec:mosaic}
+	\subsection{Puzzle} % (fold)
+	\label{sub:puzzle}
+	Our problem is to generate a minimal mosaic puzzle.
+	Mosaic\footnote{Also known as: fill-a-pix} is a relatively simple puzzle.
+	You are given a grid of squares in which some cells have a number.
+	You should color each cell either white or black.
+	A number indicates how many cells in the surrounding $3\times3$ should be colored black.
+	Below you can find an example of a mosaic puzzle and its solution.
+
+	\begin{figure}[H]
+		\centering
+		\includesvg[width=0.8\textwidth]{resources/mosaic}
+	\end{figure}
+
+	The puzzle we generate should satisfy the following requirements:
+	\begin{enumerate}
+		\item The puzzle has an unique solution.
+		\item This solution should match the solution we provide as input.
+		\item There are no other puzzles with the same solution that have fewer numbers given.
+	\end{enumerate}
+
+	We have made two programs to generate puzzles.
+	The first uses a SMT solver iteratively to find constraints that are necessary for a puzzle to be unique.
+Then it runs another SMT solver to minimize the number of hints.
+The other approach uses the unsat core.
+While Z3 finds a small puzzle using this method, it is not guaranteed to satisfy (3).
+	% subsection puzzle (end)
+
+	\subsection{Generation using unsat core} % (fold)
+	\label{sub:mosaic_unsat_core}
+	Since we want to use the unsat core, we need to make a set of constraints that is not satisfiable.
+We will use this unsatisfiability to guarantee requirement (1), namely we will try to find a second solution.
+
+	For this we add a variable for each cell, whether it is black in the second solution or not. Since we will be adding these later, we encode this as a 0-1 constrained Z3 integer:
+
+	$$\forall_r,c:  0 \le \text{filled}_{r,c} \le 1$$
+
+	Then we will add the constraint that at least one of those variable does not match the given solution (grid), so that we force Z3 to look for a second solution:
+
+	$$\exists_{r,c}:  \text{filled}_{r,c} \ne \text{grid}_{r,c}$$
+
+	We will ask Z3 to find this second solution to the puzzle with a number in every cell.
+Since we have a desired solution, we can determine which value the numbers should have.
+For every cell we add the constraint that it has the same number of black cells surrounding it in both solutions:
+
+	$$\forall_{r,c}: \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{grid}_{r',c'} = \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{filled}_{r',c'}$$
+	This formula does not handle the edges, but we do handle those cases in our code.
+Since we are interested in the positions of the numbers in the puzzle, we track these constraints so that Z3 will give them as part of the unsat core.
+
+	Now that we have all constraints we ask Z3 to check if there is a solution.
+	If there is a solution we print this alternate solution.
+	There exist desired solutions for which this happens, but it is rare.
+
+	Otherwise we ask Z3 for an unsat core.
+This gives us a list of the number constraints that guarantee there is no second solution.
+Since these constraints correspond with numbers in the grid, we have found a puzzle that satisfies requirements (1) and (2).
+We enabled the \texttt{core.minimize} flag of Z3 to try find a minimal puzzle.
+However the minimal unsat core of Z3 is "best-effort", and there is no guarantee that it is optimal.
+In particular, the example puzzle above has been generated with this program and has 15 numbers.
+However, our other program has found the following puzzle with 13 hints:
+
+	\begin{figure}[H]
+		\centering
+		\includesvg[width=0.8\textwidth]{resources/mosaic-recursive}
+	\end{figure}
+	% subsection mosaic_unsat_core (end)
+
+	\subsection{Generation using iteration} % (fold)
+	\label{sub:mosaic_iteration}
+	In this approach we use two Z3 optimizers.
+	The first optimizer searches for constraints on where to place numbers and makes these constraints as small as possible.
+	The second optimizer tries to find the smallest placement of numbers that satisfies the constraints generated by the first optimizer.\\
+
+	The idea of the first optimizer that we try to find a puzzle that satisfies all the generated constraints but still has a different solution.
+Then we consider which hints can differentiate both solutions.
+These are exactly the cells where the count of surrounding black squares differ.
+	Any puzzle with the desired solution as it's only solution must then have a number in at least one of those squares.\\
+
+	The constraints we add to the first optimizer largely correspond with what we did for the unsat core. However, we now have three types of variables:
+
+	\begin{tabular}{ll}
+		hinted & a puzzle that satisfies the generated constraints\\
+		filled & an alternate solution to that puzzle\\
+		contributes & places where the count of surrounding black squares differ.
+	\end{tabular}
+
+	Again we encode these variables with a 0-1 constrained integer in Z3:
+
+	$$\forall_{r,c}: 0\le \text{hinted}_{r,c} \le 1$$
+	$$\forall_{r,c}: 0\le \text{filled}_{r,c} \le 1$$
+	$$\forall_{r,c}: 0\le \text{contributes}_{r,c} \le 1$$
+
+	We replace the number constraint with a new one that takes into account whether a number is placed into a cell:
+
+	$$\forall_{r,c}: \text{hinted}_{r,c}\to\sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{grid}_{r',c'} = \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{filled}_{r',c'}$$
+
+	As before we require that the alternate solution differs in at least one cell from the desired solution:
+
+	$$\exists_{r,c}:  \text{filled}_{r,c} \ne \text{grid}_{r,c}$$
+
+	Next we encode that a cell differentiates both solutions when the black counts differ:
+
+	$$\forall_{r,c}: \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{grid}_{r',c'} = \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{filled}_{r',c'}\to \text{contributes}_{r,c}=1$$
+
+	The clause we generate has a literal for every cell that distinguishes both solutions.
+Since SMT solvers are generally faster on smaller clauses, we minimize the number of cells that distinguish both solutions.
+
+	$$\sum_{r,c}\text{constributes}_{r,c} = \min_{}(\sum_{r,c}\text{contributes}_{r,c}
+	)$$
+
+	When we run the first optimizer, we can have two results.
+	If the solver returns unsat, there is no puzzle that satisfies all generated constraints and has multiple solutions.
+In other words, all puzzles satisfying the generated constraints have a unique solution.
+We then use the second optimizer to find the smallest puzzle satisfying all generated constraints.\\
+
+	In the other case the solver provides a model that includes a puzzle, an alternate solution and the places where the black counts differ.
+We will add a new clause to both optimizers that encodes that a puzzle should distinguish both solutions:
+
+	$$\exists_{r,c : \text{model}(\text{contributes}_{r,c})=1}: \text{hinted}_{r,c}$$
+
+	After adding this constraint we run the first optimizer again to see if there are still alternate solutions.\\
+
+	The second optimizer is a lot easier.
+	It only uses the hinted variables, and the constraint that these are 0 or 1.
+	It also includes all generated constraints to guarantee that the resulting puzzle has an unique solution.
+	Finally it has the goal to minimize the number of cells that are hinted.\\
+
+	This optimizer will always give a model\footnote{Except in the rare cases where even the puzzle with a number in every cell has multiple solutions}.
+	This model corresponds to a minimal puzzle with a unique solution.
+Suppose there is another puzzle that has fewer hints.
+Then this puzzle does not satisfy at least one of the generated constraints.
+But then the alternate solution we found in the step that generated this constraint is also an alternate solution to the puzzle.
+Therefore there are no smaller puzzles with a unique solution.
+	% subsection mosaic_iteration (end)
+
+	\subsection{Results} % (fold)
+	\label{sub:results}
+	We have discussed two approaches to generating minimal mosaic puzzles for a given solution.
+	The unsat-core approach is more efficient since it uses features of Z3 directly.
+We were able to generate puzzles with a $20\times20$ grid within a couple of seconds.
+Unfortunately there is no guarantee that the generated puzzle is minimal, as we have shown with the example above.\\
+
+	The iterative approach is quite slow.
+It took multiple minutes to find the puzzle that disproved minimality of the unsat-core puzzle.
+The largest puzzle we have been able to generate is $8\times 8$, and took a couple of hours of computation.
+We are however confident that it generates globally minimal puzzles.
+An added benefit of this approach is that it defines the space of unique puzzles, so we can easily adapt the goal to find unique puzzles with some other property.
+For example, we can look for a minimal puzzle that has no adjacent hints, something we cannot encode in the unsat-core approach.
+	% subsection results (end)
+
+% section mosaïc (end)
+
 \appendix

### 2022-12-15 part2/mosaic-iterative.py

@@ -1,0 +1,128 @@
+from z3 import *
+import random
+
+WIDTH = 6
+HEIGHT = 6
+
+grid = [
+	[1, 1, 0, 0, 1, 0],
+	[1, 0, 1, 0, 0, 0],
+	[1, 1, 0, 1, 0, 1],
+	[0, 1, 1, 1, 0, 1],
+	[0, 0, 1, 1, 1, 1],
+	[0, 0, 1, 1, 0, 1],
+]
+# grid = [[random.randint(0,1) for c in range(WIDTH)] for r in range(HEIGHT)]
+value = [[0 for c in range(WIDTH)] for r in range(HEIGHT)]
+
+hinted = [[Int("(h {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
+contributes = [[Int("(c {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
+filled = [[Int("(f {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
+
+solver = Optimize()
+puzzle = Optimize()
+
+numConstraint = 0
+
+numContributes = 0
+numDifferent = 0
+numHint = 0
+minHint = 0
+for r in range(HEIGHT):
+	for c in range(WIDTH):
+		solver.add(0 <= hinted[r][c])
+		solver.add(hinted[r][c] <= 1)
+		solver.add(0 <= filled[r][c])
+		solver.add(filled[r][c] <= 1)
+		solver.add(0 <= contributes[r][c])
+		solver.add(contributes[r][c] <= 1)
+
+		numContributes = contributes[r][c] + numContributes
+		numHint = hinted[r][c] + numHint
+
+		if grid[r][c] == 0:
+			numDifferent = filled[r][c] + numDifferent
+		else:
+			numDifferent = 1 - filled[r][c] + numDifferent
+
+		actual = 0
+		for r2 in range(max(r-1,0), min(r+2,HEIGHT)):
+			for c2 in range(max(c-1,0), min(c+2,WIDTH)):
+				value[r][c] = grid[r2][c2] + value[r][c]
+				actual = filled[r2][c2] + actual
+		solver.add(Or(value[r][c] == actual, hinted[r][c] == 0))
+		solver.add(Or(value[r][c] == actual, contributes[r][c] == 1))
+
+		puzzle.add(0 <= hinted[r][c])
+		puzzle.add(hinted[r][c] <= 1)
+
[…]

### 2022-12-15 part2/mosaic.py

@@ -1,0 +1,72 @@
+from z3 import *
+import random
+
+WIDTH = 6
+HEIGHT = 6
+
+grid = [
+	[1, 1, 0, 0, 1, 0],
+	[1, 0, 1, 0, 0, 0],
+	[1, 1, 0, 1, 0, 1],
+	[0, 1, 1, 1, 0, 1],
+	[0, 0, 1, 1, 1, 1],
+	[0, 0, 1, 1, 0, 1],
+]
+# grid = [[random.randint(0,1) for c in range(WIDTH)] for r in range(HEIGHT)]
+value = [[0 for c in range(WIDTH)] for r in range(HEIGHT)]
+filled = [[Int("(f {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
+
+for r in range(HEIGHT):
+	for c in range(WIDTH):
+		print(grid[r][c], end=' ')
+	print()
+print()
+
+solver = Solver()
+solver.set(':core.minimize', True)
+
+numConstraint = 0
+
+numDifferent = 0
+for r in range(HEIGHT):
+	for c in range(WIDTH):
+		solver.add(0 <= filled[r][c])
+		solver.add(filled[r][c] <= 1)
+
+		if grid[r][c] == 0:
+			numDifferent += filled[r][c]
+		else:
+			numDifferent += 1-filled[r][c]
+
+		actual = 0
+		for r2 in range(max(r-1,0), min(r+2,HEIGHT)):
+			for c2 in range(max(c-1,0), min(c+2,WIDTH)):
+				value[r][c] += grid[r2][c2]
+				actual += filled[r2][c2]
+		solver.assert_and_track(value[r][c] == actual, '{} {} {}'.format(r,c,value[r][c]))
+
+solver.add(0 < numDifferent)
+satisfiable = solver.check()
+print(satisfiable)
+if satisfiable == unsat:
+	core = solver.unsat_core()
+	print(len(core))
+	puzzle = [['.' for c in range(WIDTH)] for r in range(HEIGHT)]
+	for entry in core:
+		r, c, value = map(int, entry.__str__().split())
+		puzzle[r][c] = str(value)
+	for r in range(HEIGHT):
+		for c in range(WIDTH):
+			print(puzzle[r][c], end=' ')
+		print()
+	print()
+	quit()
+
+if satisfiable == sat:
+	model = solver.model()
+	for r in range(HEIGHT):
+		for c in range(WIDTH):
[…]

### 2022-12-15 part2/report/paper.tex
Add mpo docs

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{fancyhdr}
@@
 
+\pagebreak
 \section{Robot} % (fold)
@@
 
-	On grid (a) our program finds that the minimal solution uses 14 steps. The assignment of directions is visualized in figure \ref{robot-grid-a-solution}
-
-	On grid (b) our program also finds a minimal solution of 14 steps, which is visualised in figure \ref{robot-grid-b-solution}
-
-	On grid (c) our program also finds a minimal solution of 18 steps, which is visualised in figure \ref{robot-grid-c-solution}
+	On grid (a) our program finds that the minimal solution uses 14 steps. The
+	assignment of directions is visualized in figure
+	\ref{robot-grid-a-solution}. On grid (b) our program also finds a minimal
+	solution of 14 steps, which is visualised in figure
+	\ref{robot-grid-b-solution}. On grid (c) our program also finds a minimal
+	solution of 18 steps, which is visualised in figure
+	\ref{robot-grid-c-solution}.
 
@@
 
-% section robot (end)
+\pagebreak
+\section{Multiset Path Ordering}
+\subsection{Explanation}
+In this problem we are given a set of inequalities. These inequalities contain
+terms which are either variables or functions over other terms. Our task is to
+find a Multiset Path Ordering (MPO), which is an ordering on terms generated
+from a total quasi-ordering relation $\trianglerighteq$ on the function
+symbols, using multiset comparison.
+
+To make our lives a bit easier we created a parser which can parse the
+inequalities. To this end we created four data structures which can represent:
+terms, functions, variables and inequalities. Notably, every one of these data
+structures has a function \texttt{subterms} which recursively goes down the
+term and gives all of its decendents. This is useful for when we define the
+relations on terms.
+
+We define the relations $\trianglerighteq, \rhd, \equiv$ on functions by
+creating a new boolean variable for every relation and every pair of functions,
+i.e. $\mathtt{\trianglerighteq_{fg}, \rhd_{fg}, \equiv_{fg}}$. To get the
+correct behavior we add the constraints which define these relations, see
+below. Note, that the relation $\leq$ is purely to prevent duplicates, such as
+$\mathtt{f \trianglerighteq g \lor g \trianglerighteq f}$ \textit{and}
+$\mathtt{g \trianglerighteq f \lor f \trianglerighteq g}$.
+
+$$\mathtt{\forall_f \forall_{f \leq g} \trianglerighteq_{fg} \lor
+\trianglerighteq_{gf}}$$
+
+$$\mathtt{\forall_f \forall_{f \leq g} \equiv_{fg}\ =\ \trianglerighteq_{fg}
+\land \trianglerighteq_{gf}}$$
+
+$$\mathtt{\forall_f \trianglerighteq_{ff}}$$
+
+$$\mathtt{\forall_f \forall_{g}\ \rhd_{fg} =\
+\trianglerighteq_{fg} \land \neg \trianglerighteq_{gf}}$$
+
+$$\mathtt{\forall_f \forall_{g} \forall_{h}\ \trianglerighteq_{fg} \land
+\trianglerighteq_{gh} \to\ \trianglerighteq_{fh}}$$
+
+We define the toprelations $\succsim, \succ, \approx$ on terms by creating a new
+variable for every relation and every pair of terms, i.e. $\succsim_{st},
+\succ_{st}, \approx_{st}$. To keep everything comprehensible we subdivide the
+relations $\succ, \approx$ and create new variables $\succ_{a,st},
+\succ_{b,st}, \succ_{c,st}, \approx_{a,st}, \approx_{b,st}$. The toprelations
+are now defined using the constraints shown below.
+
+$$\forall_s \forall_t \succsim_{st}\ =\ \succ_{st} \lor \approx_{st}$$
+$$\forall_s \forall_t \succ_{st}\ =\ \succ_{a,st} \lor \succ_{b,st} \lor \succ_{c,st}$$
+$$\forall_s \forall_t \approx_{st}\ =\ \approx_{a,st} \lor \approx_{b,st}$$
+
+We define constraints for when the types do not match. For example, when
+either $s$ or $t$ is a function we restrict $\neg\approx_{a,st}$. We do the
+same for all the other subrelations, these constraints follow trivially and
+will not be written out. The constraints for when the types do match are
+nontrivial. The simple subrelation constraints are as follows.
+
+$$\forall_{s = \mathtt{f}(s_1, \dots, s_n)} \forall_t \succ_{a,st}\ =
+\bigvee_{s_i} \succsim_{s_i,t}$$
+
+$$\forall_{s = \mathtt{f}(s_1, \dots, s_n)} \forall_{t = \mathtt{g}(t_1, \dots,
+t_m)} \succ_{b,st}\ = \rhd_{\mathtt{f,g}} \land \bigwedge_{t_i} \succ_{s,t_i}$$
+
+$$\forall_s \forall_t \approx_{a,st}\ =\ \approx_{st}$$
+
+We are now left with the two constraints which use the multiset properties. We
+will first take a look at $\approx_{b,st}$. Note, when $s$ or $t$ is not a
+function, or the number of arguments in $s$ and $t$ are not the same we have
+$\neg \approx_{b,st}$. For all of the other cases, where $s=\mathtt{f}(s_1,
+\dots, t_n)$ and $t=\mathtt{g}(t_1, \dots, t_n)$, we check whether there exists
+a permutation of the arguments of $t$ such that $\forall_n \approx_{s_i,t_i}$.
+We define $\mathcal{P}(t_1, \dots, t_n)$ to give all the possible permutations
+and use $t^p_i$ to denote $t_i$ in permutation $p \in \mathcal{P}$.
+
+$$\forall_{s = \mathtt{f}(s_1, \dots, s_n)} \forall_{t = \mathtt{g}(t_1, \dots,
+t_m)} \approx_{b,st}\ =\ \equiv_{\mathtt{f,g}} \land \bigvee_{p \in
+\mathcal{P}(t_1, \dots, t_n)} \bigwedge_i \approx_{s_i,t^p_i}$$
+
+The final constraint $\succ_{c,st}$ is the most elaborate of all. We use the
+tip provided in the assignment. We generate all the possible $\mathtt{EQ}$ and
+$\varphi$, such that $\exists_{i \in \{1, \dots, n\}}\ \mathtt{EQ}(i) = \bot$
+and $\forall_i\ \mathtt{EQ}(\varphi(i)) = \top \to \forall_{j \neq i}\
+\varphi(i) \neq \varphi(j)$. We can do this by generating all $\mathtt{EQ}$ and
+$\varphi$, then filtering out the functions which do not satisfy the given
+constraints. We define $\mathcal{P}^\mathtt{EQ}_\varphi(n, m)$ to give all the
+$\mathtt{EQ}, \varphi$ pairs which satisfy the given constraints. The final
+constraint is given below, seperated for when $\mathtt{EQ}(\varphi(i)) = \bot$
+or $\mathtt{EQ}(\varphi(i)) = \top$.
+
+$$\forall_{s = \mathtt{f}(s_1, \dots, s_n)} \forall_{t = \mathtt{g}(t_1, \dots,
+t_m)} \succ_{c,st}\ =\ \equiv_{\mathtt{f,g}} \land \bigvee_{\mathtt{EQ},
+\varphi \in \mathcal{P}^\mathtt{EQ}_\varphi(n, m)}
+\left(
+	\bigwedge_{i,\mathtt{EQ}(\varphi(i)) = \bot} \succ_{s_{\varphi(i)},t_i}
+	\bigwedge_{i,\mathtt{EQ}(\varphi(i)) = \top} \approx_{s_{\varphi(i)},t_i}
+\right)
+$$
+\subsection{Results}
+Running Z3 with all the constraints and variable definitions as given above
+provides us with the following ordering: $\mathtt{a \rhd c \rhd f \rhd g \rhd b
+\rhd h \rhd d }$. Finding another symbol ordering is rather easy, we just added
+the constraint $\mathtt{\neg \rhd_{h,d}}$. This gives us a new symbol ordering:
+$\mathtt{a \rhd c \rhd b \equiv f \rhd g \rhd d \rhd h}$.
+
+The human-readable proof for the inequality
+$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z)) \rhd
+\mathtt{f}(\mathtt{d}(x,z),u)$ given the first ordering is as follows.
+By 2b it suffices to proof that $\mathtt{h \rhd d}$ which holds, and
+$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~x$, and
+$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~z$. Now
+these two inequalities trivially follow from recursively applying rule 2a until
+a variable $x$ or $z$ is reached inside the arguments of $\mathtt{h}$, using 3a
+on the variables completes the proof.
 

### 2022-12-15 part2/multiset-path-order.py

@@ -11,12 +11,9 @@ class Term:
 	def print(self) -> None:
 		print(self.toString())
 		return
-	
+
 	def subterms(self):
 		return [self]
-
-	def children(self):
-		return []
 
 	def __eq__(self, other):
 		return self.toString() == other.toString()
@@ -45,15 +42,12 @@ class Function(Term):
 	def toString(self, print_index=True) -> str:
 		argStrings = [arg.toString(print_index) for arg in self.args]
 		return self.name + '(' + ', '.join(argStrings) + ')'
-	
+
 	def subterms(self):
 		terms = [self]
 		for arg in self.args:
 			terms += arg.subterms()
 		return terms
-
-	def children(self):
-		return self.args
 
 class Variable(Term):
 	name : str
@@ -118,7 +112,6 @@ inequalities.append(parse_inequality('h(g(x,g(u,z)),c(x,y,x,z)) > f(d(x,z),u)',3
 inequalities.append(parse_inequality('h(d(f(x,y),g(u,v)),f(x,y)) > f(c(u,x,v,y),g(y,x))',4))
 inequalities.append(parse_inequality('f(b(x,y,z),u) > h(u,f(x,h(y,x)))',5))
 inequalities.append(parse_inequality('b(a(x,y,z),y,x) > c(x,x,y,x)',6))
-# inequalities.append(parse_inequality('f(x) > x',7))
 
 solver = Solver()
 solver.set(':core.minimize', True)
@@ -180,8 +173,7 @@ for ineq in inequalities:
 				solver.add(Not(term_compare['>'][s][t]))
 
 			if isinstance(s, Function):
-				# solver.add(term_compare['>'][s][t] == Or(term_compare['>a'][s][t], term_compare['>b'][s][t], term_compare['>c'][s][t]))
-				solver.add(term_compare['>'][s][t] == Or(term_compare['>a'][s][t]))
+				solver.add(term_compare['>'][s][t] == Or(term_compare['>a'][s][t], term_compare['>b'][s][t], term_compare['>c'][s][t]))
 				solver.add(term_compare['>a'][s][t] == Or(*[term_compare['>~'][si][t] for si in s.args]))
 
 				if isinstance(t, Function):
@@ -234,23 +226,23 @@ for ineq in inequalities:
 
 	solver.add(term_compare['>'][ineq.lhs][ineq.rhs])
 
-# solver.add(Not(fun_ge['h']['d']))
 
-low = 0
[…]

### 2022-12-22 part2/multiset-path-order.py
Add comments

@@ -32,6 +32,7 @@ class Term:
 
 
 class Function(Term):
+	"""A class representing a function term with its arguments"""
 	name : str
 	args : list[Term]
 
@@ -49,7 +50,9 @@ class Function(Term):
 			terms += arg.subterms()
 		return terms
 
+
 class Variable(Term):
+	"""A class representing a variable."""
 	name : str
 	index : int
 
@@ -66,6 +69,7 @@ class Variable(Term):
 	def subterms(self):
 		return [self]
 
+
 def parse_at_pos(string : str, index : int, pos : int) -> (Term, int):
 	while pos < len(string) and not string[pos].isalpha():
 		pos += 1
@@ -81,11 +85,14 @@ def parse_at_pos(string : str, index : int, pos : int) -> (Term, int):
 		args.append(arg)
 	return Function(name, *args), pos+1
 
+
 def parse(string : str, index : int) -> Term:
-	t, pos = parse_at_pos(string, index, 0)
+	t, _pos = parse_at_pos(string, index, 0)
 	return t
 
+
 class Comparison:
+	"""A class representing an inequality."""
 	lhs : Term
 	rhs : Term
 
@@ -93,50 +100,58 @@ class Comparison:
 		self.lhs = lhs
 		self.rhs = rhs
 
+	def subterms(self):
+		"""Returns the unique subterms of the lhs and rhs in sorted order."""
+		return sorted(list(set(self.lhs.subterms()))), sorted(list(set(self.rhs.subterms())))
+
 	def toString(self):
 		return self.lhs.toString() + ' > ' + self.rhs.toString()
 
 	def print(self):
 		print(self.toString())
 
-def parse_inequality(string : str, index : int) -> Comparison:
+
+uid = 0  # used to generate unique variable names
+def parse_inequality(string : str) -> Comparison:
+	global uid
 	lhs, rhs = string.split('>')
-	lhs = parse(lhs, index)
-	rhs = parse(rhs, index)
+	lhs = parse(lhs, uid)
+	rhs = parse(rhs, uid)
+	uid += 1
 	return Comparison(lhs, rhs)
 
-inequalities = []
-inequalities.append(parse_inequality('c(x,y,u,v) > b(f(x,y),b(u,u,u),g(v,b(x,y,u)))',1))
-inequalities.append(parse_inequality('b(f(x,y),g(x,y),f(x,g(z,u))) > b(f(x,z),y,g(g(g(y,x),x),x))',2))
[…]

### 2022-12-22 part2/mosaic-iterative.py
Renaming variables is mosaic-iterative

@@ -16,7 +16,7 @@ grid = [
 value = [[0 for c in range(WIDTH)] for r in range(HEIGHT)]
 
 hinted = [[Int("(h {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
-contributes = [[Int("(c {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
+differs = [[Int("(c {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
 filled = [[Int("(f {},{})".format(r,c)) for c in range(WIDTH)] for r in range(HEIGHT)]
 
 solver = Optimize()
@@ -24,8 +24,8 @@ puzzle = Optimize()
 
 numConstraint = 0
 
-numContributes = 0
 numDifferent = 0
+numChanged = 0
 numHint = 0
 minHint = 0
 for r in range(HEIGHT):
@@ -34,16 +34,16 @@ for r in range(HEIGHT):
 		solver.add(hinted[r][c] <= 1)
 		solver.add(0 <= filled[r][c])
 		solver.add(filled[r][c] <= 1)
-		solver.add(0 <= contributes[r][c])
-		solver.add(contributes[r][c] <= 1)
+		solver.add(0 <= differs[r][c])
+		solver.add(differs[r][c] <= 1)
 
-		numContributes = contributes[r][c] + numContributes
+		numDifferent = differs[r][c] + numDifferent
 		numHint = hinted[r][c] + numHint
 
 		if grid[r][c] == 0:
-			numDifferent = filled[r][c] + numDifferent
+			numChanged = filled[r][c] + numChanged
 		else:
-			numDifferent = 1 - filled[r][c] + numDifferent
+			numChanged = 1 - filled[r][c] + numChanged
 
 		actual = 0
 		for r2 in range(max(r-1,0), min(r+2,HEIGHT)):
@@ -51,21 +51,21 @@ for r in range(HEIGHT):
 				value[r][c] = grid[r2][c2] + value[r][c]
 				actual = filled[r2][c2] + actual
 		solver.add(Or(value[r][c] == actual, hinted[r][c] == 0))
-		solver.add(Or(value[r][c] == actual, contributes[r][c] == 1))
+		solver.add(Or(value[r][c] == actual, differs[r][c] == 1))
 
 		puzzle.add(0 <= hinted[r][c])
 		puzzle.add(hinted[r][c] <= 1)
 
[…]

### 2022-12-22 part2/multiset-path-order.py
Fix the mpo code

@@ -147,18 +147,18 @@ functions = [term.name for term in subterms if isinstance(term, Function)]
 functions = sorted(list(set(functions)))
 
 # Generate the function ordering variables
-fun_ge = {f : {g : Bool(f + '|>=' + g) for g in functions} for f in functions}
-fun_gt = {f : {g : Bool(f + '|>' + g) for g in functions} for f in functions}
-fun_eq = {f : {g : Bool(f + '=' + g) for g in functions} for f in functions}
+fun_ge = {f : {g : Bool(f + ' |>= ' + g) for g in functions} for f in functions}
+fun_gt = {f : {g : Bool(f + ' |> ' + g) for g in functions} for f in functions}
+fun_eq = {f : {g : Bool(f + ' = ' + g) for g in functions} for f in functions}
 
 # Generate the function ordering constraints
 for f in functions:
 	for g in functions:
 		if f <= g:
 			solver.add(Or(fun_ge[f][g], fun_ge[g][f]))
-			solver.add(fun_eq[f][g] == And(fun_ge[f][g], fun_ge[g][f]))
 		for h in functions:
 			solver.add(Implies(And(fun_ge[f][g], fun_ge[g][h]), fun_ge[f][h]))
+		solver.add(fun_eq[f][g] == And(fun_ge[f][g], fun_ge[g][f]))
 		solver.add(fun_gt[f][g] == And(fun_ge[f][g], Not(fun_ge[g][f])))
 	solver.add(fun_ge[f][f])
 
@@ -175,16 +175,18 @@ for symbol in symbols:
 		for s in left_terms:
 			term_compare[symbol][s] = {}
 			for t in right_terms:
-				term_compare[symbol][s][t] = Bool(s.toString() + symbol + t.toString())
+				term_compare[symbol][s][t] = Bool(s.toString() + ' ' + symbol + ' ' + t.toString())
 
 # The total number of comparisons made with >, used to simplify the proof
 compares = []
 
-for ineq in inequalities:
+for index, ineq in enumerate(inequalities):
 	left_terms, right_terms = ineq.subterms()
 	for s in left_terms:
 		for t in right_terms:
-			compares.append(term_compare['>'][s][t])
+			if index == 2:
+				for symbol in symbols:
+					compares.append(term_compare[symbol][s][t])
 
 			# Set constraint 1
[…]

### 2022-12-22 part2/report/paper.tex
Update report

@@
 $$
+
 \subsection{Results}
-Running Z3 with all the constraints and variable definitions as given above
-provides us with the following ordering: $\mathtt{a \rhd c \rhd f \rhd g \rhd b
-\rhd h \rhd d }$. Finding another symbol ordering is rather easy, we just added
-the constraint $\mathtt{\neg \rhd_{h,d}}$. This gives us a new symbol ordering:
-$\mathtt{a \rhd c \rhd b \equiv f \rhd g \rhd d \rhd h}$.
+Running Z3 with all the constraints and variable definitions as given above provides us with the following order: $\mathtt{a \rhd c \equiv d \rhd b \rhd f \equiv h \rhd g}$.
+For this order we get the following proof for inequality 3:\\
+\begin{tabular}{l|lll}
+	Line & Statement & Rule & Children\\
+	\hline
+	1 & $u \approx u$ & 3a & \\
+	2 & $u\succsim u$ & 1 & 1\\
+	3 & $x \approx x$ & 3a & \\
+	4 & $z \approx z$ & 3a & \\
+	5 & $g(u,z) \succ u$ & 2a & 2\\
+	6 & $g(u,z) \succsim u$ & 1 & 5\\
+	7 & $g(x, g(u, z)) \succ u$ & 2a & 6\\
+	8 & $c(x, y, x, z) \succ d(x,z)$ & 2c & $c\equiv d$, 3, 4, $A_2=\{\{x,y\}\}$\\
+	9 & $h(g(x, g(u, z)), c(x, y, x, z)) \succ f(d(x, z), u)$ & 2c & $h \equiv f$, 7, 8
+\end{tabular}\\
+This is the only order for which all inequalities are provable.
 
-The human-readable proof for the inequality
-$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z)) \rhd
-\mathtt{f}(\mathtt{d}(x,z),u)$ given the first ordering is as follows.
-By 2b it suffices to proof that $\mathtt{h \rhd d}$ which holds, and
-$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~x$, and
-$\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~z$. Now
-these two inequalities trivially follow from recursively applying rule 2a until
-a variable $x$ or $z$ is reached inside the arguments of $\mathtt{h}$, using 3a
-on the variables completes the proof.
+% Running Z3 with all the constraints and variable definitions as given above
+% provides us with the following ordering: $\mathtt{a \rhd c \rhd f \rhd g \rhd b
+% \rhd h \rhd d }$. Finding another symbol ordering is rather easy, we just added
+% the constraint $\mathtt{\neg \rhd_{h,d}}$. This gives us a new symbol ordering:
+% $\mathtt{a \rhd c \rhd b \equiv f \rhd g \rhd d \rhd h}$.
+
+% The human-readable proof for the inequality
+% $\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z)) \rhd
+% \mathtt{f}(\mathtt{d}(x,z),u)$ given the first ordering is as follows.
+% By 2b it suffices to proof that $\mathtt{h \rhd d}$ which holds, and
+% $\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~x$, and
+% $\mathtt{h}(\mathtt{g}(x,\mathtt{g}(u,z)),\mathtt{c}(x,y,x,z))~\succ~z$. Now
+% these two inequalities trivially follow from recursively applying rule 2a until
+% a variable $x$ or $z$ is reached inside the arguments of $\mathtt{h}$, using 3a
+% on the variables completes the proof.
 
@@
 		filled & an alternate solution to that puzzle\\
-		contributes & places where the count of surrounding black squares differ.
+		differs & places where the count of surrounding black squares differ.
 	\end{tabular}
@@
 	$$\forall_{r,c}: 0\le \text{filled}_{r,c} \le 1$$
-	$$\forall_{r,c}: 0\le \text{contributes}_{r,c} \le 1$$
+	$$\forall_{r,c}: 0\le \text{differs}_{r,c} \le 1$$
 
@@
 
-	$$\forall_{r,c}: \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{grid}_{r',c'} = \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{filled}_{r',c'}\to \text{contributes}_{r,c}=1$$
+	$$\forall_{r,c}: \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{grid}_{r',c'} = \sum_{\substack{r-1\le r' \le r+1\\c-1\le c'\le c+1}}\text{filled}_{r',c'}\to \text{differs}_{r,c}=1$$
 
@@
 
-	$$\sum_{r,c}\text{constributes}_{r,c} = \min_{}(\sum_{r,c}\text{contributes}_{r,c}
+	$$\sum_{r,c}\text{differs}_{r,c} = \min_{}(\sum_{r,c}\text{differs}_{r,c}
 	)$$
@@
 
-	$$\exists_{r,c : \text{model}(\text{contributes}_{r,c})=1}: \text{hinted}_{r,c}$$
+	$$\exists_{r,c : \text{model}(\text{differs}_{r,c})=1}: \text{hinted}_{r,c}$$
 
